---
title: "Descriptives"
author: "Paul Stricker"
date: "2023-11-21"
output: 
   html_document:
      keep_md: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r include=FALSE}
library(dplyr)
library(tidyr)
library(patchwork)
library(ggplot2)
library(gt)
library(readxl)
library(knitr)
library(cowplot)
library(lattice)
library(extrafont)
library(xtable)
library(sf)
library(rnaturalearth)
library(rnaturalearthdata)

loadfonts(device = "win")

load("port_data.RData")
load("port_data_TEU.RData")

port_data <- filter(port_data, aggregate==0, stat_port==1) 
port_data_TEU <- filter(port_data_TEU, aggregate==0, stat_port==1)


# add control variables 
wpi_data <- read.csv("WPI_data.csv")

port_data_TEU <- merge(port_data_TEU, wpi_data, by.x = "port_code", by.y="port_code", all.x = T)
```

# What is the relevant sample? 

## Which type of cargo?

Most throughput of energy commodities (transported in bulks) and Ro-Ro comes from countries other than China. The highest share of Chinese products is shipped with container cargo. Therefore, the sample is restricted to container cargo only.

```{r , echo=FALSE,message=FALSE}
theme_set(theme_minimal() + theme(legend.position = 'none'))

port_data %>% mutate(china = ((sender == "Hong Kong" | sender == "China (except Hong Kong)"))) %>%
  pivot_longer(cols = c("roro", "bulk", "container", "other"), names_to = "type", values_to = "throughput") %>%
  group_by(type, china) %>% summarize(throughput = sum(throughput, na.rm = T)) %>% 
  ggplot(aes(y=throughput, x=reorder(type, desc(throughput)))) + geom_bar(stat="identity", position = "stack", aes(fill = china) ) +
  xlab("cargo type") + 
  labs(title="Throughput by type", caption = "Blue bars represent cargo from China.")


```

## Container weight versus volume 

```{r echo=FALSE, warning=FALSE}
merge(port_data, port_data_TEU, by.x = c("port_code", "sender_code", "year", "quarter"), by.y = c("port_code", "sender_code", "year", "quarter")) %>% 
  select(container, total_TEU, empty_TEU) %>% 
  mutate(loaded_TEU = total_TEU-empty_TEU) %>%
  ggplot(aes(x=loaded_TEU, y=container)) + geom_point() +
  ylab("container throughput (thousand tonnes)") +
         xlab("container throughput (thousand TEU)") +
         labs(title = "Container throughput: Weight versus volume (2000-2022)", caption = "Container throughput by TEU only includes loaded containers.") 


```

Container processing can be measured with two variables: Container weight or volume (TEU). Above graph shows, that the relation between the two is mostly linear. To make the results of our analysis more comparable to other studies and data sources, we will proceed using data on container TEUs. 


## Which ports are relevant? 

```{r include=FALSE}
through_put_by_ports <- port_data_TEU %>% filter(year>=2010) %>% mutate(china = ((sender == "Hong Kong" | sender == "China (except Hong Kong)") &                                          total_TEU>0)) %>%
                        group_by(port) %>% summarise(throughput=sum(total_TEU, na.rm=T), china=max(china), 
                                                     nat_stat_group = max(nat_stat_group),  
                                                     average_yrly=max(TEU_yrly)) %>% ungroup() %>%
                        mutate(throughput = throughput/sum(throughput)) %>% arrange(desc(throughput)) %>% 
                        mutate(cumsum=cumsum(throughput))
                        
```

Most of the throughput concentrates in a few hundred ports. Limiting the sample to `r sum(through_put_by_ports$cumsum<.99)` ports that are responsible for 99 percent of the container shipments might be a sensible option (`r sum(through_put_by_ports$nat_stat_group[through_put_by_ports$cumsum<.99])` of these are aggregates of ports). Additionally, not all ports process Chinese goods. While all major ports do, especially small ports usually do not. 

```{r echo=FALSE}

distribution_ports <- port_data_TEU %>% filter(year>=2010) %>% mutate(china = ((sender == "Hong Kong" | sender == "China (except Hong Kong)") & total_TEU>0)) %>%
         group_by(port) %>% summarise(throughput=sum(total_TEU, na.rm=T), china=max(china, na.rm = T)) %>% ungroup() %>%
         mutate(throughput = throughput/sum(throughput)) %>% arrange(desc(throughput)) %>% mutate(n=row_number()) %>%
         ggplot( aes(x=n, y=cumsum(throughput))) + geom_point(aes(colour = china)) + 
         ylab("share of throughput") +
         xlab("number of ports") +
         #labs(title = "Cumulative container throughput by ports (2010-2023)") +
         theme_minimal() +
              theme(panel.background = element_blank(),
                    strip.text = element_text(hjust = 0, size=12),
                    text = element_text(size=14, family="LM Roman 10"),
                    legend.position = "none")

note <- ggdraw() + geom_text(aes(label="Note: Blue dots are ports that process Chinese containers. \nData source: Eurostat."), x = 0.05, y = 0, hjust = 0,vjust = -1, size = 5, family="LM Roman 10")

cum_thrpt <- plot_grid(distribution_ports, note, ncol = 1, rel_heights = c(0.85,0.15))
cum_thrpt
ggsave(width=12, height=7, filename = "cumulative_throughput_share.pdf", path = file.path(getwd(), "figures"))



```

```{r, include=F}
cut_offs <- port_data_TEU %>% filter(year>=2010) %>% mutate(china = ((sender == "Hong Kong" | sender == "China (except Hong Kong)") &                                          total_TEU>0)) %>%
                        group_by(port) %>% summarise(throughput=sum(total_TEU, na.rm=T), china=max(china), 
                                                     nat_stat_group = max(nat_stat_group),  
                                                     average_yrly = max(TEU_yrly)) %>% ungroup() %>%
                        mutate(throughput = throughput/sum(throughput)) %>% arrange(average_yrly) %>% 
                        mutate(cumsum=cumsum(throughput))

```


An alternative to filtering by the aggregated throughput could be filtering by the average throughput. The graph below display the cumulative total throughput relative to the average yearly throughput. The graph indicates that possible cut-off thresholds of 10, 20, and 30 thousand TEUs of yearly average container throughput are associated with a loss of `r round(max(cut_offs$cumsum[cut_offs$average_yrly<=10000])*100, digits = 1 )`, `r round(max(cut_offs$cumsum[cut_offs$average_yrly<=20000])*100, digits = 1 )`, and `r round(max(cut_offs$cumsum[cut_offs$average_yrly<=30000])*100, digits = 1 )` percent of the cumulative total container throughput. In the following, the sample is constructed by applying a threshold of 20 thousand TEUs of average yearly container throughput.  

```{r echo=FALSE,message=FALSE}


port_data_TEU %>% filter(year>=2010) %>% mutate(china = ((sender == "Hong Kong" | sender == "China (except Hong Kong)") & total_TEU>0)) %>%
         group_by(port) %>% summarise(throughput=sum(total_TEU, na.rm=T), 
                                      china=max(china, na.rm = T),
                                      average_yrly = mean(TEU_yrly, na.rm=T)) %>% ungroup() %>%
         mutate(throughput = throughput/sum(throughput)) %>% arrange(average_yrly) %>% 
         ggplot( aes(x=average_yrly, y=cumsum(throughput))) + geom_point(aes(colour = china)) + 
         geom_vline(xintercept = 10000) +
         geom_vline(xintercept = 20000) +
         geom_vline(xintercept = 30000) +
         coord_cartesian(ylim=c(0,0.04), xlim=c(0,50000)) +
         ylab("cummulative share of throughput") +
         xlab("average yearly container throughput") +
         labs(title = "Possible cut-offs", caption = "Blue dots are ports that process chinese goods.")


```


```{r, include=FALSE}
port_data_TEU <- filter(port_data_TEU, TEU_yrly >=20000, port_code!="DEHAM") #exclude port that have an average yearly container throughput above 20000

port_data_agg <- port_data_TEU %>% group_by(port, port_code, period) %>% 
                               summarize(container = sum(total_TEU, na.rm=T),
                                         group = max(group),
                                         year = max(year),
                                         ownership_china = max(ownership_china),
                                         lat = mean(lat),
                                         long = mean(long)) %>% ungroup() %>%
                              group_by(port_code) %>% 
                              mutate(port_id = cur_group_id()) %>% ungroup()

#exclude Rotterdam
port_data_TEU$total_TEU[port_data_TEU$port_code=="NLRTM" & between(port_data_TEU$year,2011,2020)]<-NA
port_data_agg$container[port_data_agg$port_code=="NLRTM" & between(port_data_agg$year,2011,2020)]<-NA

```

# Sample

## Descriptives

```{r include=FALSE}

# share of national statistical groups
share <- port_data_TEU %>% mutate(stat_port = stat_port*total_TEU,
                              nat_stat_group = nat_stat_group*total_TEU) %>%
                       summarize(stat_port = sum(stat_port, na.rm=T),
                                 nat_stat_group = sum(nat_stat_group, na.rm = T)) %>%
                       mutate(share = round(nat_stat_group/(stat_port)*100, digits = 1)) %>%
                       pull(share)

#number of reporting countries
n_reporter <- as.numeric(length(unique(port_data_TEU$reporter[!is.na(port_data_TEU$total_TEU)])))


# number of ports
n_ports <- as.numeric(length(unique(port_data_TEU$port[!is.na(port_data_TEU$total_TEU)])))

# number of sending countries
n_sender <- as.numeric(length(unique(port_data_TEU$sender[!is.na(port_data_TEU$total_TEU)])))

# years covered
start_year <- min(port_data_TEU$year)
end_year <- max(port_data_TEU$year)

# number of periods
n_quarters <- as.numeric(length(unique(port_data_TEU$time[!is.na(port_data_TEU$total_TEU)])))

# number of ports with zero throughput
n_ports_nz <- port_data_TEU %>% group_by(port) %>% 
              summarise(throughput = sum(total_TEU)) %>% 
              filter(throughput==0) %>% nrow() %>%
              as.numeric()

```


The statistical units in the Eurostat data are a mixed list of actual ports and grouped ports (statistical ports) and geographic areas, which represent aggregates of many ports. While the port groups "usually" have one port management, this cannot be assumed for everyone. In addition, statistical ports that represent a group of ports make up the majority of the container shipments (`r share` percent). The definition of the statistical units (statistical ports and geographic aggregates) can be found in the [metadata](https://eur-lex.europa.eu/legal-content/EN/TXT/?uri=uriserv:OJ.L_.2018.180.01.0029.01.ENG&toc=OJ:L:2018:180:FULL), and in the [annex](https://ec.europa.eu/eurostat/cache/metadata/Annexes/mar_esms_an_2.xlsx). As of now, the sample is restricted to statistical ports (no geographic aggregates). The remaining sample includes:

- `r n_reporter` reporting countries
- `r n_ports` ports 
- `r n_sender` origin countries
- `r end_year-start_year` years (`r start_year`-`r end_year`, covering `r n_quarters` quarters)

```{r include=F}
nat_stat_groups <- read_xlsx("mar_esms_an_2.xlsx", sheet = "1.  2024 extended list of ports")  %>% 
                 rename(port = `Country/Port Name`, stat_port = `Statistical Port`, nat_stat_group = `Nat. Stat. Group`) %>%
                 filter(!is.na(nat_stat_group)) %>% select(nat_stat_group) %>% group_by(nat_stat_group) %>% summarize(count=n())

nat_stat_names <- read_xlsx("mar_esms_an_2.xlsx", sheet = "1.  2024 extended list of ports")  %>% 
                  rename(port = `Country/Port Name`, stat_port = `Statistical Port`) %>% 
                  select(port, UNLocode)

nat_stat_groups <- merge(nat_stat_groups, nat_stat_names, by.x = "nat_stat_group", by.y = "UNLocode", all.x = T)

through_put_2019 <- port_data_TEU %>% filter(year==2019) %>% group_by(port) %>%  summarize(container_2019=sum(total_TEU,na.rm=T))

nat_stat_groups <- merge(nat_stat_groups, through_put_2019, by.x="port", by.y = "port", all.x = T) %>% arrange(desc(count)) %>% 
  mutate(share = round(container_2019/count, digits = 0))

na_count_nat_stat_g <- sum(is.na(nat_stat_groups))

nat_stat_groups <- filter(nat_stat_groups, !is.na(container_2019))

```

```{r echo=FALSE}
# kable(nat_stat_groups,
#       caption = "National Statistical Groups combining many Ports")
```



## Distribution of throughput by ports and origin



```{r include=FALSE}
# top 50 ports
top50 <- port_data_TEU %>% filter(year>2010) %>% group_by(port) %>% summarise(throughput=sum(total_TEU, na.rm=T)) %>% ungroup() %>% 
         arrange(desc(throughput)) %>% head(n=50) %>% pull(port)

# top 50 ports with highest share of chinese goods
top50_china <- port_data_TEU %>% filter(year>2010) %>% mutate(china = ((sender == "Hong Kong" | sender == "China (except Hong Kong)") & total_TEU>0),
                                                    throughput_others = case_when(china==1 ~ 0,
                                                                                  china==0 ~ total_TEU),
                                                    throughput_china = case_when(china==1 ~ total_TEU,
                                                                                 china==0 ~ 0)) %>%
         group_by(port) %>% summarise(throughput_others=sum(throughput_others, na.rm=T),
                                      throughput_china=sum(throughput_china, na.rm=T),
                                      share = throughput_china/throughput_others) %>% arrange(desc(share)) %>% 
         head(n=50) %>% pull(port)

```

Port that process a relatively high share of cargo from China are generally large ports. Below are displayed the top 50 largest ports, their aggregate throughput and their throughput of Chinese container shipments (Mainland China + Hong Kong). The other graph displays the top 50 ports with the highest share of Chinese container throughput. Both lists are rather similar with `r as.numeric(length(setdiff(top50,top50_china)))` large ports which are not among the top 50 ports with high Chinese container throughput. 

```{r echo=FALSE, warning=FALSE, message=FALSE}
port_data_TEU %>% filter(year>=2010, port %in% top50) %>%  mutate(china = ((sender == "Hong Kong" | sender == "China (except Hong Kong)") & total_TEU>0)) %>%
              group_by(port, china) %>% summarise(throughput=sum(total_TEU, na.rm=T)) %>% ungroup() %>%  
              ggplot(aes(x=reorder(port,desc(throughput)), y=throughput, fill=china)) + 
              geom_bar(stat="identity", position = "stack") +
              labs(title="Container throughput of the 50 largest ports (2010-2023)", caption = "Blue bars represent throughput of chinese goods.")+
              theme(axis.text.x = element_text(angle = 45, hjust = 1),
                    axis.title.x = element_blank())

port_data_TEU %>% filter(year>=2010, port %in% top50_china) %>%  mutate(china = ((sender == "Hong Kong" | sender == "China (except Hong Kong)") & total_TEU>0)) %>%
              group_by(port, china) %>% summarise(throughput=sum(total_TEU, na.rm=T)) %>% ungroup() %>%  
              ggplot(aes(x=reorder(port,desc(throughput)), y=throughput, fill=china)) + 
              geom_bar(stat="identity", position = "stack") +
              labs(title="Container throughput of the 50 ports with the highest share of chinese throughput (2010-2023)", caption = "Blue bars represent throughput of chinese goods.")+
              theme(axis.text.x = element_text(angle = 45, hjust = 1),
                    axis.title.x = element_blank())
```

```{r include=FALSE}
#china share
china_share <- port_data_TEU %>% mutate(china = ((sender == "Hong Kong" | sender == "China (except Hong Kong)") & total_TEU >0),
                                        throughput_others = case_when(china==1 ~ 0,
                                                                      china==0 ~ total_TEU),
                                        throughput_china = case_when(china==1 ~ total_TEU,
                                                                     china==0 ~ 0)) %>%
  group_by(year) %>% summarise(throughput_others=sum(throughput_others, na.rm=T),
                               throughput_china=sum(throughput_china, na.rm=T),
                               share = throughput_china/throughput_others*100*2500000)

```
Above graphs indicate that the share of Chinese cargo is (surprisingly) low. The graph below plots the total container shipments from China, the total container shipments from origins other than China and the share of Chinese containers in percent. Recent global crises, like the financial crisis in 2007-2009, the corona-pandemic in 2020, and the Ukraine war in 2023 have led to drops in total cargo shipments. Interestingly, despite the narrative of the ever increasing dependency on Chinese goods, the share of Chinese container shipments is on a declining path since 2012. 

```{r echo=FALSE,message=FALSE}
port_data_TEU   %>% mutate(china = ((sender == "Hong Kong" | sender == "China (except Hong Kong)") & total_TEU>0)) %>%
                group_by(year, china) %>% summarise(throughput=sum(total_TEU, na.rm=T)) %>% ungroup() %>%
                ggplot(aes(x=year)) + 
                geom_line(aes(y=throughput, color = china)) +
                geom_line(data=china_share, aes(x=year,y=share)) + 
                scale_y_continuous(
                  name = "throughput",
                  sec.axis = sec_axis( trans=~./2500000, name="share in percent")) + 
                labs(title = "Aggregate container throughput over time", 
                caption = "Blue line represents throughput of chinese goods. 
                           Red line represents throughput from all other countries.
                           Black line represents share in percent.") +
                theme(axis.title.y.right = element_text(angle = 90))

```



```{r echo=FALSE, message=FALSE}

port_data_TEU %>% mutate(china = ((sender == "Hong Kong" | sender == "China (except Hong Kong)") & total_TEU>0),
                     sender = case_when(china==0 ~ sender,
                                        china==1 ~ "China")) %>%
              group_by(sender) %>% summarize(throughput=sum(total_TEU, na.rm=T), china=max(china)) %>% ungroup() %>% 
              arrange(desc(throughput)) %>% head(n=50) %>%
              ggplot(aes(x=reorder(sender, desc(throughput)), y=throughput)) + 
              geom_bar(stat="identity", aes(fill=china)) +
              labs(title = "Top 50 origins of container shipments (2010-2023)") +
              theme(axis.text.x = element_text(angle = 45, hjust = 1),
                    axis.title.x = element_blank())
```

## Chinese ownership treatment

```{r include=FALSE}

length(unique(port_data_TEU$port_code[port_data_TEU$group!=0]))

```

Originally, we found 26 ports which were at some point at least partly owned by a Chinese company. The selected sample contains `r length(unique(port_data_TEU$port_code[port_data_TEU$group!=0]))` treated ports. Eight ports dropped out of the sample due to various reasons:

- One port does not report any data (Willebroek)
- One falls into an aggregate of many ports (Venlo, is included in an aggregate of 96 ports)
- One falls below the average throughput threshold of 20 million TEUs (Stockholm)
- Four ports were always treated (Duisburg, Felixtowe, Thamesport, Harwich)
- On is excluded due to being treated only very recently (Hamburg)

The following graph display all 18 treated ports and their annual container throughput as a share of their mean container throughput. The vertical lines indicate treatment time. Solid line represent the container throughput of the respective port. The dashed line represents the average throughput of the control group. Note that not all ports report the throughput for every period. This might influence the average.

```{r echo=FALSE, message=FALSE, warning=F, fig.dim = c(8, 14)}
treated_ports <- port_data_TEU %>% filter(group!=0) %>% 
                 group_by(port_code, time) %>% 
                 summarize(total_TEU = sum(total_TEU, na.rm=T)) %>% 
                 group_by(port_code) %>% summarise(max = max(total_TEU, na.rm=T)) %>% 
                 arrange(desc(max)) %>% pull(port_code)

plot_list <- list()

for (i in 1:length(treated_ports)){
  treatment_period <- unique(port_data_TEU$group[port_data_TEU$port_code==treated_ports[i]])
  treatment_time <- unique(port_data_TEU$time[port_data_TEU$period==treatment_period])
  port_name <- unique(port_data_TEU$port[port_data_TEU$port_code==treated_ports[i]])
  
  plot_list[[i]] <- port_data_TEU %>% filter(port_code==treated_ports[i] | group==0) %>% group_by(port_code, time) %>%
                                         summarize(sum = sum(total_TEU, na.rm=T),
                                                   ownership_china = max(ownership_china),
                                                   group = max(group),
                                                   year = max(year)) %>% 
                                                  #group_by(port_code) %>% 
                                                  #mutate(sum = sum / mean(sum, na.rm=T)) %>% 
                                                   group_by(group, time) %>% 
                                         summarize(count = n_distinct(port_code),
                                                   container = ifelse(port_code=="NLRTM" & between(year,2011,2020), NA, sum(sum, na.rm=T)/count),
                                                   ownership_china = max(ownership_china)) %>%
              ggplot(aes(x=time, y=container/1000)) + geom_line(aes(linetype=as.factor(group))) +
              geom_vline(xintercept = treatment_time) +
              scale_linetype_manual(values=c("dotted", "solid")) +
              labs(title = port_name ) +
              theme_minimal() +
              theme(panel.background = element_blank(),
                    strip.text = element_text(hjust = 0, size=12),
                    text = element_text(size=10, family="LM Roman 10"),
                    legend.position = "none",
                    axis.title.x = element_blank(),
                    axis.title.y = element_blank())
}

x.title <- ggdraw() + geom_text(aes(label="TEU (millions)", angle = 90), x = 0.5, y = 0.5, hjust = 0.5, vjust = 1, size = 5, family="LM Roman 10")
y.title <- ggdraw() + geom_text(aes(label="Year"), x = 0.5, y = 0.5, hjust = 0.5, vjust = 1, size = 5, family="LM Roman 10")

note <- ggdraw() + geom_text(aes(label="Note: Vertical line indicates treatment timing."), x = 0.05, y = 0, hjust = 0,vjust = -1, size = 3, family="LM Roman 10")

legend <- get_legend(plot_list[[1]] + 
                       scale_linetype_discrete(labels=c( "treated", "average of untreated")) +
                       theme(legend.position="bottom", legend.title = element_blank()))

plots <- plot_grid(plotlist = plot_list, ncol=3, align = "hv") 

plots1 <- plot_grid(plots, y.title, legend,note, ncol=1, rel_heights = c(0.925,0.025, 0.025, 0.025) ) 
#plots1 <- plot_grid(plots1, note, ncol=1, rel_heights = c(0.975,0.025) ) 

plot_grid(x.title, plots1, ncol = 2, rel_widths = c(0.05,0.95))

ggsave(width=10, height=15, filename = "port_ts_graph.pdf", path = file.path(getwd(), "figures"))
```

```{r include=F}

# calculate summary statistics by group

#controls
sumstat_controls <- port_data_agg %>% group_by(port_id) %>% mutate(treated=max(ownership_china)) %>% ungroup() %>%
                                      filter(treated==0, container!=0) %>% arrange(port_id, period) %>%
                                      group_by(port_id) %>%
                                      summarize(across(container, list(agrate = ~ (tail(.x,n=1)-head(.x,n=1))/(tail(period,n=1)-head(period,n=1)))),
                                                across(container, ~ mean(.x, na.rm=T), .names="{col}_mean")) %>%
                                      ungroup() %>% select(-port_id) %>%
                                      summarize(across(ends_with("mean"), mean, .names = "{col}"),
                                                across(ends_with("grate"),~ mean(.x,na.rm=T), .names = "{col}")) %>% 
                                      pivot_longer(cols = starts_with(c("container_")), 
                                                   names_to = c("vars", ".value"),
                                                   names_pattern = "(.*)_(.*)")
#treated 
sumstat_treated <- port_data_agg %>% group_by(port_id) %>% mutate(treated=max(ownership_china))  %>%
                                      filter(treated==1, container!=0) %>% arrange(port_id, period) %>%
                                      mutate(across(container, ~ mean(.x, na.rm=T), .names="{col}_mean")) %>%
                                      group_by(port_id, ownership_china) %>%
                                      summarize(across(container, list(agrate = ~ (tail(.x,n=1)-head(.x,n=1))/(tail(period,n=1)-head(period,n=1)))),
                                                across(container_mean, ~ max(.x))) %>%
                                      group_by(ownership_china) %>% 
                                      summarize(across(ends_with("mean"), mean, .names = "{col}"),
                                                across(ends_with("grate"),~ mean(.x,na.rm=T), .names = "{col}")) %>%
                                      pivot_wider(names_from = "ownership_china", values_from = ends_with("grate"), names_glue = "{.value}{ownership_china}") %>% 
                                      pivot_longer(cols = starts_with(c("container_")), 
                                                   names_to = c("vars", ".value"),
                                                   names_pattern = "(.*)_(.*)")

#combine
sumstats <- merge(sumstat_controls, sumstat_treated, by=c("vars")) %>%
            mutate(across(c("mean.x","agrate","mean.y","agrate0","agrate1"), ~ format(round(.,digits=0), big.mark = ",")))
  
    

```

```{r include=TRUE}
# generate latex table

# Insert multirow cell into the appropriate position in the LaTeX code
addtorow <- list()
addtorow$pos <- list(nrow(sumstats))
addtorow$command <- c("[2ex] \\hline \\hline \\multicolumn{6}{l}{\\shortstack[l]{\\small Note: $\\Delta$ refers to average quarterly changes. Ports recieved equal \\\\ weights in the calculation of means and average changes.}}\\\\")

# Modify the LaTeX code
latex_code <- print.xtable(xtable(sumstats, caption = "Summary statistics"),
                           label = "tab:sumstats",
                           caption.placement = "top",
                           include.rownames = FALSE,
                           add.to.row = addtorow,
                           hline.after = c(-1,-1,0),
                           size = "\\begin{adjustbox}{max width=\\textwidth}")

latex_code <- gsub("\\\\end\\{tabular\\}", "\\\\end{tabular} \n \\\\end{adjustbox}", latex_code)
latex_code <- gsub("vars & mean.x & agrate & mean.y & agrate0 & agrate1"," & \\\\multicolumn{2}{c}{Control group [n=XX]} & \\\\multicolumn{3}{c}{Treatment group [n=XX]} \\\\\\\\\n \\\\cmidrule(lr){2-3} \\\\cmidrule(lr){4-6}  Var. & mean & $\\\\Delta$ & mean & $\\\\Delta$ [$t \\\\leq t^*$] & $\\\\Delta$ [$t>t^*$]", latex_code)

# Save modified LaTeX code to a .tex file
cat(latex_code, file = file.path(getwd(), "results_tables", "sumstats.tex"))


# generate a gt table
sumstats_gt <- gt(sumstats, rowname_col = "vars") |>
              tab_header(
                title = "Summary statistics") |>
              cols_label(
                mean.x="mean",
                agrate="av. delta",
                mean.y="mean",
                agrate0="av.delta (t<t*)",
                agrate1="av.delta (t>t*)"
              ) |>
              tab_spanner(label = "Control group",
                          columns = c("mean.x", "agrate")) |>
              tab_spanner(label = "Treated",
                          columns = c("mean.y", "agrate0", "agrate1")) |>
              sub_missing(columns = everything(),
                          rows = everything(),
                          missing_text = "") |>
              tab_footnote(footnote = "") |>
              tab_options(stub.border.width = 0,
                          table_body.hlines.width = 0
              )

sumstats_gt


```

```{r include=FALSE}
# generate map 
european_ports <- port_data_agg %>% filter(year==2020 | year==2000, !is.na(lat), !is.na(long)) %>% group_by(port, year) %>%
                                    summarize(container = sum(container),
                                              lat = mean(lat),
                                              long = mean(long),
                                              ownership_china = max(ownership_china)) %>% group_by(port) %>%
                                    mutate(ownership_china=max(ownership_china)) %>%
                                    pivot_wider(names_from = year, values_from = container, names_prefix = "y") %>%
                                    mutate(delta=y2020-y2000,
                                           growth=(y2020/y2000-1)*100)


world <- ne_countries(scale = "medium", returnclass = "sf")    

# Create a ggplot object
ggmap_change <- ggplot(data=world) + geom_sf() +
            geom_point(data = european_ports[european_ports$ownership_china==0,], aes(x = long, y = lat, size = delta, color = ownership_china), shape = 24, color="blue", fill="blue") +
            geom_point(data = european_ports[european_ports$ownership_china==1,], aes(x = long, y = lat, size = delta), shape = 24, color = "red", fill="red") + #shape 24 would be trianlges
            coord_sf(xlim = c(-13,28), ylim = c(34,60), expand = FALSE)  + 
            theme(#legend.position = "none",
                  axis.title = element_blank(),
                  axis.text = element_blank(),
                  axis.ticks.x = element_blank(),
                  axis.ticks.length = unit(.1, "pt"),
                  axis.ticks.y = element_blank(),
                  panel.border = element_rect(color = "black", fill = NA, size = 1.5),
                  panel.grid.major = element_blank(),
                  panel.grid.minor = element_blank(),
                  panel.background = element_rect(fill = "grey98"),
                  plot.margin = margin(0, 0, 0, 0)
                  )+
  scale_fill_gradient(na.value="white",low = "grey80", high = "grey30")






```

```{r warning=F}
# plot

ggmap_change

```


## Data availability for treated ports

The DiD approach is sensible to missing data in the timeseries of treated units. In the best case, there are no missing values in each treated unit. There are two options to achieve this: First, trim the time horizon at the front an end to make sure all treated ports start and end at the same time. Second, missing values in the middle of the time series could be imputed. 

```{r echo=FALSE, message=FALSE, fig.dim = c(16,8)}
missingness_matrix <- port_data_agg %>% filter(group!=0) %>% mutate(missing=case_when(!is.na(container) & ownership_china==0 ~ 1,
                                                                                      !is.na(container) & ownership_china==1 ~2)) %>%
                        pivot_wider(id_cols = "period", names_from = "port", values_from = "missing", values_fill = 0) %>%
                        arrange(period) %>% select(-period)
missingness_matrix <- as.matrix(missingness_matrix)

levelplot(missingness_matrix)
```


