---
title: "Port investments as policy tools"
author: "Katja Kalkschmied"
date: "2024-07-09"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Difference in Differences: cross-port variation in trade with ports in Lithuania

```{r include=FALSE, warning=FALSE, message=FALSE}
library(plm)
library(dplyr)
library(tidyr)
library(stringr)
library(patchwork)
library(ggplot2)
library(fixest)
library(did)
library(CInf)
library(readxl)
library(gt)
library(xtable)
library(extrafont)



loadfonts(device = "win")
set.seed(100)

#load data
load("port_data_TEU_109.RData")

port_data_TEU <- filter(port_data_TEU, aggregate==0,
                                       stat_port==1,              #filter out non-stat ports (and geographic aggregates)
                                       TEU_yrly >=20000,          #filter out small ports
                                       port_code!="DEHAM",        #exclude Hamburg due to very recent treatment
                                       sender_code=="LT") %>%      #include only containers to and from Lithuania
             mutate(s_china = ifelse(sender_code=="CN_X_HK" | sender_code=="HK", 1, 0)) %>%
             group_by(port_code, sender_code) %>%
             mutate(pair_id=cur_group_id()) %>% ungroup() %>%
             rename(container=total_TEU)


port_data_agg <- port_data_TEU %>% mutate(origin_china = ifelse(!is.na(container) & (sender_code=="CN_X_HK" | sender_code=="HK"),container,0)) %>%
                               group_by(port, port_code, period, reporter_code) %>% 
                               summarize(container = sum(container, na.rm=T),
                                         real_gdp_reporter = mean(real_gdp_reporter),
                                         origin_china = sum(origin_china),
                                         group = max(group),
                                         group_operation = max(group_operation),
                                         ownership_china = max(ownership_china),
                                         operation = max(operation),
                                         year = max(year),
                                         quarter = max(quarter)) %>% ungroup() %>%
                              group_by(port_code) %>% 
                              mutate(port_id = cur_group_id(),
                                     mean_container = mean(container, na.rm=T)) %>% ungroup()

# manipulate group variable to set treatment to September 2021 = period 100
port_data_agg <- port_data_agg %>%
  mutate(group = ifelse(group == 0, 0, 100))


#exclude Rotterdam
port_data_TEU$container[port_data_TEU$port_code=="NLRTM" & between(port_data_TEU$year,2011,2020)]<-NA
port_data_agg$container[port_data_agg$port_code=="NLRTM" & between(port_data_agg$year,2011,2020)]<-NA

port_data_agg_ny <- port_data_agg %>% filter(group!=0)

port_list <- port_data_agg %>% filter(group!=0) %>% group_by(port) %>% summarize(id = mean(port_id))

port_group_list <- port_data_agg %>% filter(group!=0) %>% group_by(group) %>%
                                     summarize(ports = toString(unique(port)))
port_group_list$ports <- paste(port_group_list$ports,paste("(", port_group_list$group, ")", sep = ""))

# Add Control variables --------------------------------------------------------

# add terminal data to the data set
#terminal_data <- read_xlsx("port_manual_berth.xlsx")

load("berth_data.RData")
terminal_data <- berth_data
  
# add channel depth and cranes from World port index
wpi_data <- read.csv("WPI_data.csv")

# control variables
control_var <- merge(wpi_data, terminal_data, by.x="port_code",by.y="port_code", all = T) %>%
               select(port_code, 
                      year, quarter, quay_length, cranes, 
                      cranes_fixed, cranes_mobile, cranes_floating, max_lift, cargo_depth)

# merge with aggregated data
port_data_agg <- merge(port_data_agg, control_var, by.x = c("port_code", "year", "quarter"), by.y = c("port_code", "year", "quarter"), all.x = T) 

#port_data_agg$ships[port_data_agg$ships==0] <- 1

port_data_agg$quay_length[port_data_agg$quay_length==0] <- 100
port_data_agg$cranes[port_data_agg$cranes==0] <- 1

### Control variables for DID ####


controls <- c("cranes", "max_lift", "cargo_depth")
var_controls <- c("quay_length")

#controls <- c("cranes", "max_lift", "cargo_depth")
#var_controls <- c("quay_length")

```



```{r include=F, message=FALSE, warning=FALSE}

#### Estimation ot ATT with never treated as control group ####
attgt <- simple_staggered_did(yname = "container",            #outcome variable
                              tname = "period",                 #time variable
                              idname = "port_id",                    #id for units (should be port_code*sender)
                              gname = "group",                    #grouping variable for first treatment. Is zero for never treated units)
                              #xformula = c("ships"),
                              #varformula = c("ships"),
                              #control_group = "not_yet_treated",
                              data = port_data_agg)


attgt_cond <- simple_staggered_did(yname = "container",            #outcome variable
                              tname = "period",                 #time variable
                              idname = "port_id",                    #id for units (should be port_code*sender)
                              gname = "group",                    #grouping variable for first treatment. Is zero for never treated units)
                              xformula = controls,
                              varformula = var_controls,
                              #control_group = "not_yet_treated",
                              data = port_data_agg)



# putting together the results

results1 <- data.frame(Estimate = c("simple average", port_list$port[match(as.character(attgt$groupwise_treatment_effects$id), port_list$id)]),
                       ATT = c(attgt$average_treatment_effect$att,
                                   attgt$groupwise_treatment_effects$att),
                       lower = c(attgt$average_treatment_effect$att - attgt$average_treatment_effect$crit_val,
                                attgt$groupwise_treatment_effects$att - attgt$groupwise_treatment_effects$crit_val),
                       upper = c(attgt$average_treatment_effect$att + attgt$average_treatment_effect$crit_val,
                                  attgt$groupwise_treatment_effects$att + attgt$groupwise_treatment_effects$crit_val),
                       p_value = c(NA, attgt$groupwise_treatment_effects$pre_treatment_p_value))

results2 <- data.frame(Estimate = c("simple average", port_list$port[match(as.character(attgt_cond$groupwise_treatment_effects$id), port_list$id)]),
                       ATT_cond = c(attgt_cond$average_treatment_effect$att,
                                   attgt_cond$groupwise_treatment_effects$att),
                       lower_cond = c(attgt_cond$average_treatment_effect$att - attgt_cond$average_treatment_effect$crit_val,
                                attgt_cond$groupwise_treatment_effects$att - attgt_cond$groupwise_treatment_effects$crit_val),
                       upper_cond = c(attgt_cond$average_treatment_effect$att + attgt_cond$average_treatment_effect$crit_val,
                                  attgt_cond$groupwise_treatment_effects$att + attgt_cond$groupwise_treatment_effects$crit_val),
                        p_value_cond = c(NA, attgt_cond$groupwise_treatment_effects$pre_treatment_p_value))

results <- merge(results1, results2, by.x="Estimate", by.y="Estimate", all=T,  sort=F) %>%
               mutate(sig = (ATT>0 & lower>0) | (ATT<0 & upper<0),
                      sig_cond = (ATT_cond>0 & lower_cond>0) | (ATT_cond<0 & upper_cond<0),
                      pretest = (p_value < 0.05) & !is.na(p_value),
                      pretest_cond = (p_value_cond < 0.05) & !is.na(p_value_cond),
                      across(c("ATT","ATT_cond","lower","upper","lower_cond","upper_cond"), ~ format(round(.,digits=0), big.mark = ",")),

                      ATT = ifelse(sig==T, paste0(ATT, "*"), ATT),
                      ATT = ifelse(pretest==T, paste0(ATT, "^{x}"),ATT),
                      ATT_cond = ifelse(sig_cond==T, paste0(ATT_cond, "*"), ATT_cond),
                      ATT_cond = ifelse(pretest_cond==T, paste0(ATT_cond, "^{x}"),ATT_cond),
                      Conf = paste0("[",str_trim(lower),", ",str_trim(upper),"]"),
                      Conf_cond = paste0("[",str_trim(lower_cond),"; ",str_trim(upper_cond),"]"))
results <- results[c("Estimate", "ATT", "Conf", "ATT_cond", "Conf_cond")]



#### Estimate ATT with not-yet-treated control group ####
attgt_ny <- simple_staggered_did(yname = "container",            #outcome variable
                              tname = "period",                 #time variable
                              idname = "port_id",                    #id for units (should be port_code*sender)
                              gname = "group",                    #grouping variable for first treatment. Is zero for never treated units)
                              #xformula = c("ships"),
                              #varformula = c("ships"),
                              control_group = "not_yet_treated",
                              data = port_data_agg)


attgt_cond_ny <- simple_staggered_did(yname = "container",            #outcome variable
                              tname = "period",                 #time variable
                              idname = "port_id",                    #id for units (should be port_code*sender)
                              gname = "group",                    #grouping variable for first treatment. Is zero for never treated units)
                              xformula = controls,
                              varformula = var_controls,
                              control_group = "not_yet_treated",
                              data = port_data_agg)



# putting together the results

results_ny1 <- data.frame(Estimate = c("simple average", port_list$port[match(as.character(attgt_ny$groupwise_treatment_effects$id), port_list$id)]),
                       ATT_ny = c(attgt_ny$average_treatment_effect$att,
                                   attgt_ny$groupwise_treatment_effects$att),
                       lower_ny = c(attgt_ny$average_treatment_effect$att - attgt_ny$average_treatment_effect$crit_val,
                                attgt_ny$groupwise_treatment_effects$att - attgt_ny$groupwise_treatment_effects$crit_val),
                       upper_ny = c(attgt_ny$average_treatment_effect$att + attgt_ny$average_treatment_effect$crit_val,
                                  attgt_ny$groupwise_treatment_effects$att + attgt_ny$groupwise_treatment_effects$crit_val),
                       p_value_ny = c(NA, attgt_ny$groupwise_treatment_effects$pre_treatment_p_value))

results_ny2 <- data.frame(Estimate = c("simple average", port_list$port[match(as.character(attgt_cond_ny$groupwise_treatment_effects$id), port_list$id)]),
                       ATT_cond_ny = c(attgt_cond_ny$average_treatment_effect$att,
                                   attgt_cond_ny$groupwise_treatment_effects$att),
                       lower_cond_ny = c(attgt_cond_ny$average_treatment_effect$att - attgt_cond_ny$average_treatment_effect$crit_val,
                                attgt_cond_ny$groupwise_treatment_effects$att - attgt_cond_ny$groupwise_treatment_effects$crit_val),
                       upper_cond_ny = c(attgt_cond_ny$average_treatment_effect$att + attgt_cond_ny$average_treatment_effect$crit_val,
                                  attgt_cond_ny$groupwise_treatment_effects$att + attgt_cond_ny$groupwise_treatment_effects$crit_val),
                       p_value_cond_ny = c(NA, attgt_cond_ny$groupwise_treatment_effects$pre_treatment_p_value)) 

results_ny <- merge(results_ny1, results_ny2, by.x="Estimate", by.y="Estimate", all=T,  sort=F) %>%
               mutate(sig = (ATT_ny>0 & lower_ny>0) | (ATT_ny<0 & upper_ny<0),
                      sig_cond = (ATT_cond_ny>0 & lower_cond_ny>0) | (ATT_cond_ny<0 & upper_cond_ny<0),
                      pretest = (p_value_ny < 0.05) & !is.na(p_value_ny),
                      pretest_cond = (p_value_cond_ny < 0.05) & !is.na(p_value_cond_ny),
                      across(c("ATT_ny","ATT_cond_ny","lower_ny","upper_ny","lower_cond_ny","upper_cond_ny"), ~ format(round(.,digits=0), big.mark = ",")),
                      ATT_ny = ifelse(sig==T, paste0(ATT_ny, "*"), ATT_ny),
                      ATT_ny = ifelse(pretest==T, paste0(ATT_ny, "^{x}"),ATT_ny),
                      ATT_cond_ny = ifelse(sig_cond==T, paste0(ATT_cond_ny, "*"), ATT_cond_ny),
                      ATT_cond_ny = ifelse(pretest_cond==T, paste0(ATT_cond_ny, "^{x}"),ATT_cond_ny),
                      Conf_ny = paste0("[",str_trim(lower_ny),", ",str_trim(upper_ny),"]"),
                      Conf_cond_ny = paste0("[",str_trim(lower_cond_ny),"; ",str_trim(upper_cond_ny),"]"))
results_ny <- results_ny[c("Estimate", "ATT_cond_ny", "Conf_cond_ny")]

results <- merge(results, results_ny, by.x="Estimate", by.y="Estimate", sort=F)


```

```{r include=F}
# generating a gt table
results_gt <- gt(results, rowname_col = "Estimate") |>
              tab_header(
                title = "Aggregation of group time treatment effects, treatment in period 100",
                subtitle = "Dependent variable: Total throughput with Lithuania in thousand TEU") |>
              cols_label(
                ATT="unconditional",
                ATT_cond="conditional",
                ATT_cond_ny="conditional",
                Conf="95% CI",
                Conf_cond="95% CI",
                Conf_cond_ny="95% CI"
              ) |>
              tab_spanner(label = "Never treated",
                          columns = c("ATT", "Conf", "ATT_cond", "Conf_cond")) |>
              tab_spanner(label = "Not yet treated",
                          columns = c("ATT_cond_ny", "Conf_cond_ny")) |>
              tab_row_group(label = "by group:",
                            rows = c(2:10)) |>
              row_group_order(groups = c(NA,"by group:")) |>
              sub_missing(columns = everything(),
                          rows = everything(),
                          missing_text = "") |>
              tab_footnote(footnote = "Stars indicate that the confidence interval does not include zero.") |>
              tab_options(stub.border.width = 0,
                          table_body.hlines.width = 0
              )

```

```{r include=TRUE, echo=FALSE}
results_gt

```

```{r include=TRUE, echo=FALSE}
average_container <- port_data_agg %>%
  summarise(average_container_sample = mean(container, na.rm = TRUE))

print(average_container)

average_container_gdynia <- port_data_agg %>%
  filter(port == "Gdynia") %>%
  summarise(average_container_gdynia = mean(container, na.rm = TRUE))

print(average_container_gdynia)
```

Comparing trade with Lithuanian ports of Chinese treated ports with non-Chinese treated ports after the opening of the diplomatic council with Taiwan in September 2021, we do see a small negative bias of Chinese treated ports. This, however, is solely driven by the port of Gdynia. The ports of Antwerpen and Le Havre show a positive effect. The effect of Antwerpen is suspicious/problematic since our recording for Antwerpen stops at 100. So there is only one treated data point for Antwerpen. We have to recheck what to do with this. Since we have recordings for Antwerp-Grugges from 101-107, we may consider merging them into one time series.

Gdynia is not a trivial port for Lithuania. It is geographically close and trade twice as much as the average port in our sample.

Difference to previous estimation tasks: I drew new information from Eurostat to increase the period from 107 (2023, Q3) to 109 (2024, Q1) to have more post-treatment observations. 

# Difference in Differences: check if Gdynia effect is a general effect

```{r include=FALSE, warning=FALSE, message=FALSE}

loadfonts(device = "win")
set.seed(100)

#load data
load("port_data_TEU_109.RData")

port_data_TEU <- filter(port_data_TEU, aggregate==0,
                                       stat_port==1,              #filter out non-stat ports (and geographic aggregates)
                                       TEU_yrly >=20000,          #filter out small ports
                                       port_code!="DEHAM") %>%    #exclude Hamburg due to very recent treatment
             mutate(s_china = ifelse(sender_code=="CN_X_HK" | sender_code=="HK", 1, 0)) %>%
             group_by(port_code, sender_code) %>%
             mutate(pair_id=cur_group_id()) %>% ungroup() %>%
             rename(container=total_TEU)


port_data_agg <- port_data_TEU %>% mutate(origin_china = ifelse(!is.na(container) & (sender_code=="CN_X_HK" | sender_code=="HK"),container,0)) %>%
                               group_by(port, port_code, period, reporter_code) %>% 
                               summarize(container = sum(container, na.rm=T),
                                         real_gdp_reporter = mean(real_gdp_reporter),
                                         origin_china = sum(origin_china),
                                         group = max(group),
                                         group_operation = max(group_operation),
                                         ownership_china = max(ownership_china),
                                         operation = max(operation),
                                         year = max(year),
                                         quarter = max(quarter)) %>% ungroup() %>%
                              group_by(port_code) %>% 
                              mutate(port_id = cur_group_id(),
                                     mean_container = mean(container, na.rm=T)) %>% ungroup()

# manipulate group variable to set treatment to September 2021 = period 100
port_data_agg <- port_data_agg %>%
  mutate(group = ifelse(group == 0, 0, 100))


#exclude Rotterdam
port_data_TEU$container[port_data_TEU$port_code=="NLRTM" & between(port_data_TEU$year,2011,2020)]<-NA
port_data_agg$container[port_data_agg$port_code=="NLRTM" & between(port_data_agg$year,2011,2020)]<-NA

port_data_agg_ny <- port_data_agg %>% filter(group!=0)

port_list <- port_data_agg %>% filter(group!=0) %>% group_by(port) %>% summarize(id = mean(port_id))

port_group_list <- port_data_agg %>% filter(group!=0) %>% group_by(group) %>%
                                     summarize(ports = toString(unique(port)))
port_group_list$ports <- paste(port_group_list$ports,paste("(", port_group_list$group, ")", sep = ""))

# Add Control variables --------------------------------------------------------

# add terminal data to the data set
#terminal_data <- read_xlsx("port_manual_berth.xlsx")

load("berth_data.RData")
terminal_data <- berth_data
  
# add channel depth and cranes from World port index
wpi_data <- read.csv("WPI_data.csv")

# control variables
control_var <- merge(wpi_data, terminal_data, by.x="port_code",by.y="port_code", all = T) %>%
               select(port_code, 
                      year, quarter, quay_length, cranes, 
                      cranes_fixed, cranes_mobile, cranes_floating, max_lift, cargo_depth)


# merge with aggregated data
port_data_agg <- merge(port_data_agg, control_var, by.x = c("port_code", "year", "quarter"), by.y = c("port_code", "year", "quarter"), all.x = T) 

#port_data_agg$ships[port_data_agg$ships==0] <- 1

port_data_agg$quay_length[port_data_agg$quay_length==0] <- 100
port_data_agg$cranes[port_data_agg$cranes==0] <- 1

### Control variables for DID ####


controls <- c("cranes", "max_lift", "cargo_depth")
var_controls <- c("quay_length")

#controls <- c("cranes", "max_lift", "cargo_depth")
#var_controls <- c("quay_length")

```



```{r include=F, message=FALSE, warning=FALSE}

#### Estimation ot ATT with never treated as control group ####
attgt <- simple_staggered_did(yname = "container",            #outcome variable
                              tname = "period",                 #time variable
                              idname = "port_id",                    #id for units (should be port_code*sender)
                              gname = "group",                    #grouping variable for first treatment. Is zero for never treated units)
                              #xformula = c("ships"),
                              #varformula = c("ships"),
                              #control_group = "not_yet_treated",
                              data = port_data_agg)


attgt_cond <- simple_staggered_did(yname = "container",            #outcome variable
                              tname = "period",                 #time variable
                              idname = "port_id",                    #id for units (should be port_code*sender)
                              gname = "group",                    #grouping variable for first treatment. Is zero for never treated units)
                              xformula = controls,
                              varformula = var_controls,
                              #control_group = "not_yet_treated",
                              data = port_data_agg)



# putting together the results

results1 <- data.frame(Estimate = c("simple average", port_list$port[match(as.character(attgt$groupwise_treatment_effects$id), port_list$id)]),
                       ATT = c(attgt$average_treatment_effect$att,
                                   attgt$groupwise_treatment_effects$att),
                       lower = c(attgt$average_treatment_effect$att - attgt$average_treatment_effect$crit_val,
                                attgt$groupwise_treatment_effects$att - attgt$groupwise_treatment_effects$crit_val),
                       upper = c(attgt$average_treatment_effect$att + attgt$average_treatment_effect$crit_val,
                                  attgt$groupwise_treatment_effects$att + attgt$groupwise_treatment_effects$crit_val),
                       p_value = c(NA, attgt$groupwise_treatment_effects$pre_treatment_p_value))

results2 <- data.frame(Estimate = c("simple average", port_list$port[match(as.character(attgt_cond$groupwise_treatment_effects$id), port_list$id)]),
                       ATT_cond = c(attgt_cond$average_treatment_effect$att,
                                   attgt_cond$groupwise_treatment_effects$att),
                       lower_cond = c(attgt_cond$average_treatment_effect$att - attgt_cond$average_treatment_effect$crit_val,
                                attgt_cond$groupwise_treatment_effects$att - attgt_cond$groupwise_treatment_effects$crit_val),
                       upper_cond = c(attgt_cond$average_treatment_effect$att + attgt_cond$average_treatment_effect$crit_val,
                                  attgt_cond$groupwise_treatment_effects$att + attgt_cond$groupwise_treatment_effects$crit_val),
                        p_value_cond = c(NA, attgt_cond$groupwise_treatment_effects$pre_treatment_p_value))

results <- merge(results1, results2, by.x="Estimate", by.y="Estimate", all=T,  sort=F) %>%
               mutate(sig = (ATT>0 & lower>0) | (ATT<0 & upper<0),
                      sig_cond = (ATT_cond>0 & lower_cond>0) | (ATT_cond<0 & upper_cond<0),
                      pretest = (p_value < 0.05) & !is.na(p_value),
                      pretest_cond = (p_value_cond < 0.05) & !is.na(p_value_cond),
                      across(c("ATT","ATT_cond","lower","upper","lower_cond","upper_cond"), ~ format(round(.,digits=0), big.mark = ",")),

                      ATT = ifelse(sig==T, paste0(ATT, "*"), ATT),
                      ATT = ifelse(pretest==T, paste0(ATT, "^{x}"),ATT),
                      ATT_cond = ifelse(sig_cond==T, paste0(ATT_cond, "*"), ATT_cond),
                      ATT_cond = ifelse(pretest_cond==T, paste0(ATT_cond, "^{x}"),ATT_cond),
                      Conf = paste0("[",str_trim(lower),", ",str_trim(upper),"]"),
                      Conf_cond = paste0("[",str_trim(lower_cond),"; ",str_trim(upper_cond),"]"))
results <- results[c("Estimate", "ATT", "Conf", "ATT_cond", "Conf_cond")]



#### Estimate ATT with not-yet-treated control group ####
attgt_ny <- simple_staggered_did(yname = "container",            #outcome variable
                              tname = "period",                 #time variable
                              idname = "port_id",                    #id for units (should be port_code*sender)
                              gname = "group",                    #grouping variable for first treatment. Is zero for never treated units)
                              #xformula = c("ships"),
                              #varformula = c("ships"),
                              control_group = "not_yet_treated",
                              data = port_data_agg)


attgt_cond_ny <- simple_staggered_did(yname = "container",            #outcome variable
                              tname = "period",                 #time variable
                              idname = "port_id",                    #id for units (should be port_code*sender)
                              gname = "group",                    #grouping variable for first treatment. Is zero for never treated units)
                              xformula = controls,
                              varformula = var_controls,
                              control_group = "not_yet_treated",
                              data = port_data_agg)



# putting together the results

results_ny1 <- data.frame(Estimate = c("simple average", port_list$port[match(as.character(attgt_ny$groupwise_treatment_effects$id), port_list$id)]),
                       ATT_ny = c(attgt_ny$average_treatment_effect$att,
                                   attgt_ny$groupwise_treatment_effects$att),
                       lower_ny = c(attgt_ny$average_treatment_effect$att - attgt_ny$average_treatment_effect$crit_val,
                                attgt_ny$groupwise_treatment_effects$att - attgt_ny$groupwise_treatment_effects$crit_val),
                       upper_ny = c(attgt_ny$average_treatment_effect$att + attgt_ny$average_treatment_effect$crit_val,
                                  attgt_ny$groupwise_treatment_effects$att + attgt_ny$groupwise_treatment_effects$crit_val),
                       p_value_ny = c(NA, attgt_ny$groupwise_treatment_effects$pre_treatment_p_value))

results_ny2 <- data.frame(Estimate = c("simple average", port_list$port[match(as.character(attgt_cond_ny$groupwise_treatment_effects$id), port_list$id)]),
                       ATT_cond_ny = c(attgt_cond_ny$average_treatment_effect$att,
                                   attgt_cond_ny$groupwise_treatment_effects$att),
                       lower_cond_ny = c(attgt_cond_ny$average_treatment_effect$att - attgt_cond_ny$average_treatment_effect$crit_val,
                                attgt_cond_ny$groupwise_treatment_effects$att - attgt_cond_ny$groupwise_treatment_effects$crit_val),
                       upper_cond_ny = c(attgt_cond_ny$average_treatment_effect$att + attgt_cond_ny$average_treatment_effect$crit_val,
                                  attgt_cond_ny$groupwise_treatment_effects$att + attgt_cond_ny$groupwise_treatment_effects$crit_val),
                       p_value_cond_ny = c(NA, attgt_cond_ny$groupwise_treatment_effects$pre_treatment_p_value)) 

results_ny <- merge(results_ny1, results_ny2, by.x="Estimate", by.y="Estimate", all=T,  sort=F) %>%
               mutate(sig = (ATT_ny>0 & lower_ny>0) | (ATT_ny<0 & upper_ny<0),
                      sig_cond = (ATT_cond_ny>0 & lower_cond_ny>0) | (ATT_cond_ny<0 & upper_cond_ny<0),
                      pretest = (p_value_ny < 0.05) & !is.na(p_value_ny),
                      pretest_cond = (p_value_cond_ny < 0.05) & !is.na(p_value_cond_ny),
                      across(c("ATT_ny","ATT_cond_ny","lower_ny","upper_ny","lower_cond_ny","upper_cond_ny"), ~ format(round(.,digits=0), big.mark = ",")),
                      ATT_ny = ifelse(sig==T, paste0(ATT_ny, "*"), ATT_ny),
                      ATT_ny = ifelse(pretest==T, paste0(ATT_ny, "^{x}"),ATT_ny),
                      ATT_cond_ny = ifelse(sig_cond==T, paste0(ATT_cond_ny, "*"), ATT_cond_ny),
                      ATT_cond_ny = ifelse(pretest_cond==T, paste0(ATT_cond_ny, "^{x}"),ATT_cond_ny),
                      Conf_ny = paste0("[",str_trim(lower_ny),", ",str_trim(upper_ny),"]"),
                      Conf_cond_ny = paste0("[",str_trim(lower_cond_ny),"; ",str_trim(upper_cond_ny),"]"))
results_ny <- results_ny[c("Estimate", "ATT_cond_ny", "Conf_cond_ny")]

results <- merge(results, results_ny, by.x="Estimate", by.y="Estimate", sort=F)


```

```{r include=F}
# generating a gt table
results_gt <- gt(results, rowname_col = "Estimate") |>
              tab_header(
                title = "Aggregation of group time treatment effects, treatment in period 100",
                subtitle = "Dependent variable: Total throughput with all nationality ports in thousand TEU") |>
              cols_label(
                ATT="unconditional",
                ATT_cond="conditional",
                ATT_cond_ny="conditional",
                Conf="95% CI",
                Conf_cond="95% CI",
                Conf_cond_ny="95% CI"
              ) |>
              tab_spanner(label = "Never treated",
                          columns = c("ATT", "Conf", "ATT_cond", "Conf_cond")) |>
              tab_spanner(label = "Not yet treated",
                          columns = c("ATT_cond_ny", "Conf_cond_ny")) |>
              tab_row_group(label = "by group:",
                            rows = c(2:10)) |>
              row_group_order(groups = c(NA,"by group:")) |>
              sub_missing(columns = everything(),
                          rows = everything(),
                          missing_text = "") |>
              tab_footnote(footnote = "Stars indicate that the confidence interval does not include zero.") |>
              tab_options(stub.border.width = 0,
                          table_body.hlines.width = 0
              )

```

```{r include=TRUE, echo=FALSE}
results_gt

```

When looking at changes of the treated ports with all trading partners around period 100 (September 2021) number of Chinese ports have experienced a decline in trade as compared to untreated ports. I think this is still a covid effect, that has especially hit ports engaged in international trade which are the ports in which Chinese invested. I guess international container shipments were still hurt eventually substituted by intraregional trade. 

Also, Gdynia experienced a decrease of its overall container shipments more than a comparable untreated port.

# Difference in Differences: check if general Gdynia effect at period stays if we exclude shipments to Lithuania

```{r include=FALSE, warning=FALSE, message=FALSE}

loadfonts(device = "win")
set.seed(100)

#load data
load("port_data_TEU_109.RData")

port_data_TEU <- filter(port_data_TEU, aggregate==0,
                                       stat_port==1,              #filter out non-stat ports (and geographic aggregates)
                                       TEU_yrly >=20000,          #filter out small ports
                                       port_code!="DEHAM",
                                        sender_code!="LT") %>%    #exclude Hamburg due to very recent treatment
             mutate(s_china = ifelse(sender_code=="CN_X_HK" | sender_code=="HK", 1, 0)) %>%
             group_by(port_code, sender_code) %>%
             mutate(pair_id=cur_group_id()) %>% ungroup() %>%
             rename(container=total_TEU)


port_data_agg <- port_data_TEU %>% mutate(origin_china = ifelse(!is.na(container) & (sender_code=="CN_X_HK" | sender_code=="HK"),container,0)) %>%
                               group_by(port, port_code, period, reporter_code) %>% 
                               summarize(container = sum(container, na.rm=T),
                                         real_gdp_reporter = mean(real_gdp_reporter),
                                         origin_china = sum(origin_china),
                                         group = max(group),
                                         group_operation = max(group_operation),
                                         ownership_china = max(ownership_china),
                                         operation = max(operation),
                                         year = max(year),
                                         quarter = max(quarter)) %>% ungroup() %>%
                              group_by(port_code) %>% 
                              mutate(port_id = cur_group_id(),
                                     mean_container = mean(container, na.rm=T)) %>% ungroup()

# manipulate group variable to set treatment to September 2021 = period 100
port_data_agg <- port_data_agg %>%
  mutate(group = ifelse(group == 0, 0, 100))


#exclude Rotterdam
port_data_TEU$container[port_data_TEU$port_code=="NLRTM" & between(port_data_TEU$year,2011,2020)]<-NA
port_data_agg$container[port_data_agg$port_code=="NLRTM" & between(port_data_agg$year,2011,2020)]<-NA

port_data_agg_ny <- port_data_agg %>% filter(group!=0)

port_list <- port_data_agg %>% filter(group!=0) %>% group_by(port) %>% summarize(id = mean(port_id))

port_group_list <- port_data_agg %>% filter(group!=0) %>% group_by(group) %>%
                                     summarize(ports = toString(unique(port)))
port_group_list$ports <- paste(port_group_list$ports,paste("(", port_group_list$group, ")", sep = ""))

# Add Control variables --------------------------------------------------------

# add terminal data to the data set
#terminal_data <- read_xlsx("port_manual_berth.xlsx")

load("berth_data.RData")
terminal_data <- berth_data
  
# add channel depth and cranes from World port index
wpi_data <- read.csv("WPI_data.csv")

# control variables
control_var <- merge(wpi_data, terminal_data, by.x="port_code",by.y="port_code", all = T) %>%
               select(port_code, 
                      year, quarter, quay_length, cranes, 
                      cranes_fixed, cranes_mobile, cranes_floating, max_lift, cargo_depth)


# merge with aggregated data
port_data_agg <- merge(port_data_agg, control_var, by.x = c("port_code", "year", "quarter"), by.y = c("port_code", "year", "quarter"), all.x = T) 

#port_data_agg$ships[port_data_agg$ships==0] <- 1

port_data_agg$quay_length[port_data_agg$quay_length==0] <- 100
port_data_agg$cranes[port_data_agg$cranes==0] <- 1

### Control variables for DID ####


controls <- c("cranes", "max_lift", "cargo_depth")
var_controls <- c("quay_length")

#controls <- c("cranes", "max_lift", "cargo_depth")
#var_controls <- c("quay_length")

```



```{r include=F, message=FALSE, warning=FALSE}

#### Estimation ot ATT with never treated as control group ####
attgt <- simple_staggered_did(yname = "container",            #outcome variable
                              tname = "period",                 #time variable
                              idname = "port_id",                    #id for units (should be port_code*sender)
                              gname = "group",                    #grouping variable for first treatment. Is zero for never treated units)
                              #xformula = c("ships"),
                              #varformula = c("ships"),
                              #control_group = "not_yet_treated",
                              data = port_data_agg)


attgt_cond <- simple_staggered_did(yname = "container",            #outcome variable
                              tname = "period",                 #time variable
                              idname = "port_id",                    #id for units (should be port_code*sender)
                              gname = "group",                    #grouping variable for first treatment. Is zero for never treated units)
                              xformula = controls,
                              varformula = var_controls,
                              #control_group = "not_yet_treated",
                              data = port_data_agg)



# putting together the results

results1 <- data.frame(Estimate = c("simple average", port_list$port[match(as.character(attgt$groupwise_treatment_effects$id), port_list$id)]),
                       ATT = c(attgt$average_treatment_effect$att,
                                   attgt$groupwise_treatment_effects$att),
                       lower = c(attgt$average_treatment_effect$att - attgt$average_treatment_effect$crit_val,
                                attgt$groupwise_treatment_effects$att - attgt$groupwise_treatment_effects$crit_val),
                       upper = c(attgt$average_treatment_effect$att + attgt$average_treatment_effect$crit_val,
                                  attgt$groupwise_treatment_effects$att + attgt$groupwise_treatment_effects$crit_val),
                       p_value = c(NA, attgt$groupwise_treatment_effects$pre_treatment_p_value))

results2 <- data.frame(Estimate = c("simple average", port_list$port[match(as.character(attgt_cond$groupwise_treatment_effects$id), port_list$id)]),
                       ATT_cond = c(attgt_cond$average_treatment_effect$att,
                                   attgt_cond$groupwise_treatment_effects$att),
                       lower_cond = c(attgt_cond$average_treatment_effect$att - attgt_cond$average_treatment_effect$crit_val,
                                attgt_cond$groupwise_treatment_effects$att - attgt_cond$groupwise_treatment_effects$crit_val),
                       upper_cond = c(attgt_cond$average_treatment_effect$att + attgt_cond$average_treatment_effect$crit_val,
                                  attgt_cond$groupwise_treatment_effects$att + attgt_cond$groupwise_treatment_effects$crit_val),
                        p_value_cond = c(NA, attgt_cond$groupwise_treatment_effects$pre_treatment_p_value))

results <- merge(results1, results2, by.x="Estimate", by.y="Estimate", all=T,  sort=F) %>%
               mutate(sig = (ATT>0 & lower>0) | (ATT<0 & upper<0),
                      sig_cond = (ATT_cond>0 & lower_cond>0) | (ATT_cond<0 & upper_cond<0),
                      pretest = (p_value < 0.05) & !is.na(p_value),
                      pretest_cond = (p_value_cond < 0.05) & !is.na(p_value_cond),
                      across(c("ATT","ATT_cond","lower","upper","lower_cond","upper_cond"), ~ format(round(.,digits=0), big.mark = ",")),

                      ATT = ifelse(sig==T, paste0(ATT, "*"), ATT),
                      ATT = ifelse(pretest==T, paste0(ATT, "^{x}"),ATT),
                      ATT_cond = ifelse(sig_cond==T, paste0(ATT_cond, "*"), ATT_cond),
                      ATT_cond = ifelse(pretest_cond==T, paste0(ATT_cond, "^{x}"),ATT_cond),
                      Conf = paste0("[",str_trim(lower),", ",str_trim(upper),"]"),
                      Conf_cond = paste0("[",str_trim(lower_cond),"; ",str_trim(upper_cond),"]"))
results <- results[c("Estimate", "ATT", "Conf", "ATT_cond", "Conf_cond")]



#### Estimate ATT with not-yet-treated control group ####
attgt_ny <- simple_staggered_did(yname = "container",            #outcome variable
                              tname = "period",                 #time variable
                              idname = "port_id",                    #id for units (should be port_code*sender)
                              gname = "group",                    #grouping variable for first treatment. Is zero for never treated units)
                              #xformula = c("ships"),
                              #varformula = c("ships"),
                              control_group = "not_yet_treated",
                              data = port_data_agg)


attgt_cond_ny <- simple_staggered_did(yname = "container",            #outcome variable
                              tname = "period",                 #time variable
                              idname = "port_id",                    #id for units (should be port_code*sender)
                              gname = "group",                    #grouping variable for first treatment. Is zero for never treated units)
                              xformula = controls,
                              varformula = var_controls,
                              control_group = "not_yet_treated",
                              data = port_data_agg)



# putting together the results

results_ny1 <- data.frame(Estimate = c("simple average", port_list$port[match(as.character(attgt_ny$groupwise_treatment_effects$id), port_list$id)]),
                       ATT_ny = c(attgt_ny$average_treatment_effect$att,
                                   attgt_ny$groupwise_treatment_effects$att),
                       lower_ny = c(attgt_ny$average_treatment_effect$att - attgt_ny$average_treatment_effect$crit_val,
                                attgt_ny$groupwise_treatment_effects$att - attgt_ny$groupwise_treatment_effects$crit_val),
                       upper_ny = c(attgt_ny$average_treatment_effect$att + attgt_ny$average_treatment_effect$crit_val,
                                  attgt_ny$groupwise_treatment_effects$att + attgt_ny$groupwise_treatment_effects$crit_val),
                       p_value_ny = c(NA, attgt_ny$groupwise_treatment_effects$pre_treatment_p_value))

results_ny2 <- data.frame(Estimate = c("simple average", port_list$port[match(as.character(attgt_cond_ny$groupwise_treatment_effects$id), port_list$id)]),
                       ATT_cond_ny = c(attgt_cond_ny$average_treatment_effect$att,
                                   attgt_cond_ny$groupwise_treatment_effects$att),
                       lower_cond_ny = c(attgt_cond_ny$average_treatment_effect$att - attgt_cond_ny$average_treatment_effect$crit_val,
                                attgt_cond_ny$groupwise_treatment_effects$att - attgt_cond_ny$groupwise_treatment_effects$crit_val),
                       upper_cond_ny = c(attgt_cond_ny$average_treatment_effect$att + attgt_cond_ny$average_treatment_effect$crit_val,
                                  attgt_cond_ny$groupwise_treatment_effects$att + attgt_cond_ny$groupwise_treatment_effects$crit_val),
                       p_value_cond_ny = c(NA, attgt_cond_ny$groupwise_treatment_effects$pre_treatment_p_value)) 

results_ny <- merge(results_ny1, results_ny2, by.x="Estimate", by.y="Estimate", all=T,  sort=F) %>%
               mutate(sig = (ATT_ny>0 & lower_ny>0) | (ATT_ny<0 & upper_ny<0),
                      sig_cond = (ATT_cond_ny>0 & lower_cond_ny>0) | (ATT_cond_ny<0 & upper_cond_ny<0),
                      pretest = (p_value_ny < 0.05) & !is.na(p_value_ny),
                      pretest_cond = (p_value_cond_ny < 0.05) & !is.na(p_value_cond_ny),
                      across(c("ATT_ny","ATT_cond_ny","lower_ny","upper_ny","lower_cond_ny","upper_cond_ny"), ~ format(round(.,digits=0), big.mark = ",")),
                      ATT_ny = ifelse(sig==T, paste0(ATT_ny, "*"), ATT_ny),
                      ATT_ny = ifelse(pretest==T, paste0(ATT_ny, "^{x}"),ATT_ny),
                      ATT_cond_ny = ifelse(sig_cond==T, paste0(ATT_cond_ny, "*"), ATT_cond_ny),
                      ATT_cond_ny = ifelse(pretest_cond==T, paste0(ATT_cond_ny, "^{x}"),ATT_cond_ny),
                      Conf_ny = paste0("[",str_trim(lower_ny),", ",str_trim(upper_ny),"]"),
                      Conf_cond_ny = paste0("[",str_trim(lower_cond_ny),"; ",str_trim(upper_cond_ny),"]"))
results_ny <- results_ny[c("Estimate", "ATT_cond_ny", "Conf_cond_ny")]

results <- merge(results, results_ny, by.x="Estimate", by.y="Estimate", sort=F)


```

```{r include=F}
# generating a gt table
results_gt <- gt(results, rowname_col = "Estimate") |>
              tab_header(
                title = "Aggregation of group time treatment effects, treatment in period 100",
                subtitle = "Dependent variable: Total throughput with all nationality ports but Lithuania in thousand TEU") |>
              cols_label(
                ATT="unconditional",
                ATT_cond="conditional",
                ATT_cond_ny="conditional",
                Conf="95% CI",
                Conf_cond="95% CI",
                Conf_cond_ny="95% CI"
              ) |>
              tab_spanner(label = "Never treated",
                          columns = c("ATT", "Conf", "ATT_cond", "Conf_cond")) |>
              tab_spanner(label = "Not yet treated",
                          columns = c("ATT_cond_ny", "Conf_cond_ny")) |>
              tab_row_group(label = "by group:",
                            rows = c(2:10)) |>
              row_group_order(groups = c(NA,"by group:")) |>
              sub_missing(columns = everything(),
                          rows = everything(),
                          missing_text = "") |>
              tab_footnote(footnote = "Stars indicate that the confidence interval does not include zero.") |>
              tab_options(stub.border.width = 0,
                          table_body.hlines.width = 0
              )

```

```{r include=TRUE, echo=FALSE}
results_gt

```

The negative effect for Gdynia in general trade with its partners becomes insignificant when we omit ports in Lithuania. Hence, reductions in trade with Lithuania mattered for the overall trade development of Gdynia.

# Difference in Differences: check if Gdynia effect holds for trade with Germany

```{r include=FALSE, warning=FALSE, message=FALSE}

loadfonts(device = "win")
set.seed(100)

#load data
load("port_data_TEU_109.RData")

port_data_TEU <- filter(port_data_TEU, aggregate==0,
                                       stat_port==1,              #filter out non-stat ports (and geographic aggregates)
                                       TEU_yrly >=20000,          #filter out small ports
                                       port_code!="DEHAM",        #exclude Hamburg due to very recent treatment
                                       sender_code=="DE") %>%      #include only containers to and from Lithuania
             mutate(s_china = ifelse(sender_code=="CN_X_HK" | sender_code=="HK", 1, 0)) %>%
             group_by(port_code, sender_code) %>%
             mutate(pair_id=cur_group_id()) %>% ungroup() %>%
             rename(container=total_TEU)


port_data_agg <- port_data_TEU %>% mutate(origin_china = ifelse(!is.na(container) & (sender_code=="CN_X_HK" | sender_code=="HK"),container,0)) %>%
                               group_by(port, port_code, period, reporter_code) %>% 
                               summarize(container = sum(container, na.rm=T),
                                         real_gdp_reporter = mean(real_gdp_reporter),
                                         origin_china = sum(origin_china),
                                         group = max(group),
                                         group_operation = max(group_operation),
                                         ownership_china = max(ownership_china),
                                         operation = max(operation),
                                         year = max(year),
                                         quarter = max(quarter)) %>% ungroup() %>%
                              group_by(port_code) %>% 
                              mutate(port_id = cur_group_id(),
                                     mean_container = mean(container, na.rm=T)) %>% ungroup()

# manipulate group variable to set treatment to September 2021 = period 100
port_data_agg <- port_data_agg %>%
  mutate(group = ifelse(group == 0, 0, 100))


#exclude Rotterdam
port_data_TEU$container[port_data_TEU$port_code=="NLRTM" & between(port_data_TEU$year,2011,2020)]<-NA
port_data_agg$container[port_data_agg$port_code=="NLRTM" & between(port_data_agg$year,2011,2020)]<-NA

port_data_agg_ny <- port_data_agg %>% filter(group!=0)

port_list <- port_data_agg %>% filter(group!=0) %>% group_by(port) %>% summarize(id = mean(port_id))

port_group_list <- port_data_agg %>% filter(group!=0) %>% group_by(group) %>%
                                     summarize(ports = toString(unique(port)))
port_group_list$ports <- paste(port_group_list$ports,paste("(", port_group_list$group, ")", sep = ""))

# Add Control variables --------------------------------------------------------

# add terminal data to the data set
#terminal_data <- read_xlsx("port_manual_berth.xlsx")

load("berth_data.RData")
terminal_data <- berth_data
  
# add channel depth and cranes from World port index
wpi_data <- read.csv("WPI_data.csv")

# control variables
control_var <- merge(wpi_data, terminal_data, by.x="port_code",by.y="port_code", all = T) %>%
               select(port_code, 
                      year, quarter, quay_length, cranes, 
                      cranes_fixed, cranes_mobile, cranes_floating, max_lift, cargo_depth)


# merge with aggregated data
port_data_agg <- merge(port_data_agg, control_var, by.x = c("port_code", "year", "quarter"), by.y = c("port_code", "year", "quarter"), all.x = T) 

#port_data_agg$ships[port_data_agg$ships==0] <- 1

port_data_agg$quay_length[port_data_agg$quay_length==0] <- 100
port_data_agg$cranes[port_data_agg$cranes==0] <- 1

### Control variables for DID ####


controls <- c("cranes", "max_lift", "cargo_depth")
var_controls <- c("quay_length")

#controls <- c("cranes", "max_lift", "cargo_depth")
#var_controls <- c("quay_length")

```



```{r include=F, message=FALSE, warning=FALSE}

#### Estimation ot ATT with never treated as control group ####
attgt <- simple_staggered_did(yname = "container",            #outcome variable
                              tname = "period",                 #time variable
                              idname = "port_id",                    #id for units (should be port_code*sender)
                              gname = "group",                    #grouping variable for first treatment. Is zero for never treated units)
                              #xformula = c("ships"),
                              #varformula = c("ships"),
                              #control_group = "not_yet_treated",
                              data = port_data_agg)


attgt_cond <- simple_staggered_did(yname = "container",            #outcome variable
                              tname = "period",                 #time variable
                              idname = "port_id",                    #id for units (should be port_code*sender)
                              gname = "group",                    #grouping variable for first treatment. Is zero for never treated units)
                              xformula = controls,
                              varformula = var_controls,
                              #control_group = "not_yet_treated",
                              data = port_data_agg)



# putting together the results

results1 <- data.frame(Estimate = c("simple average", port_list$port[match(as.character(attgt$groupwise_treatment_effects$id), port_list$id)]),
                       ATT = c(attgt$average_treatment_effect$att,
                                   attgt$groupwise_treatment_effects$att),
                       lower = c(attgt$average_treatment_effect$att - attgt$average_treatment_effect$crit_val,
                                attgt$groupwise_treatment_effects$att - attgt$groupwise_treatment_effects$crit_val),
                       upper = c(attgt$average_treatment_effect$att + attgt$average_treatment_effect$crit_val,
                                  attgt$groupwise_treatment_effects$att + attgt$groupwise_treatment_effects$crit_val),
                       p_value = c(NA, attgt$groupwise_treatment_effects$pre_treatment_p_value))

results2 <- data.frame(Estimate = c("simple average", port_list$port[match(as.character(attgt_cond$groupwise_treatment_effects$id), port_list$id)]),
                       ATT_cond = c(attgt_cond$average_treatment_effect$att,
                                   attgt_cond$groupwise_treatment_effects$att),
                       lower_cond = c(attgt_cond$average_treatment_effect$att - attgt_cond$average_treatment_effect$crit_val,
                                attgt_cond$groupwise_treatment_effects$att - attgt_cond$groupwise_treatment_effects$crit_val),
                       upper_cond = c(attgt_cond$average_treatment_effect$att + attgt_cond$average_treatment_effect$crit_val,
                                  attgt_cond$groupwise_treatment_effects$att + attgt_cond$groupwise_treatment_effects$crit_val),
                        p_value_cond = c(NA, attgt_cond$groupwise_treatment_effects$pre_treatment_p_value))

results <- merge(results1, results2, by.x="Estimate", by.y="Estimate", all=T,  sort=F) %>%
               mutate(sig = (ATT>0 & lower>0) | (ATT<0 & upper<0),
                      sig_cond = (ATT_cond>0 & lower_cond>0) | (ATT_cond<0 & upper_cond<0),
                      pretest = (p_value < 0.05) & !is.na(p_value),
                      pretest_cond = (p_value_cond < 0.05) & !is.na(p_value_cond),
                      across(c("ATT","ATT_cond","lower","upper","lower_cond","upper_cond"), ~ format(round(.,digits=0), big.mark = ",")),

                      ATT = ifelse(sig==T, paste0(ATT, "*"), ATT),
                      ATT = ifelse(pretest==T, paste0(ATT, "^{x}"),ATT),
                      ATT_cond = ifelse(sig_cond==T, paste0(ATT_cond, "*"), ATT_cond),
                      ATT_cond = ifelse(pretest_cond==T, paste0(ATT_cond, "^{x}"),ATT_cond),
                      Conf = paste0("[",str_trim(lower),", ",str_trim(upper),"]"),
                      Conf_cond = paste0("[",str_trim(lower_cond),"; ",str_trim(upper_cond),"]"))
results <- results[c("Estimate", "ATT", "Conf", "ATT_cond", "Conf_cond")]



#### Estimate ATT with not-yet-treated control group ####
attgt_ny <- simple_staggered_did(yname = "container",            #outcome variable
                              tname = "period",                 #time variable
                              idname = "port_id",                    #id for units (should be port_code*sender)
                              gname = "group",                    #grouping variable for first treatment. Is zero for never treated units)
                              #xformula = c("ships"),
                              #varformula = c("ships"),
                              control_group = "not_yet_treated",
                              data = port_data_agg)


attgt_cond_ny <- simple_staggered_did(yname = "container",            #outcome variable
                              tname = "period",                 #time variable
                              idname = "port_id",                    #id for units (should be port_code*sender)
                              gname = "group",                    #grouping variable for first treatment. Is zero for never treated units)
                              xformula = controls,
                              varformula = var_controls,
                              control_group = "not_yet_treated",
                              data = port_data_agg)



# putting together the results

results_ny1 <- data.frame(Estimate = c("simple average", port_list$port[match(as.character(attgt_ny$groupwise_treatment_effects$id), port_list$id)]),
                       ATT_ny = c(attgt_ny$average_treatment_effect$att,
                                   attgt_ny$groupwise_treatment_effects$att),
                       lower_ny = c(attgt_ny$average_treatment_effect$att - attgt_ny$average_treatment_effect$crit_val,
                                attgt_ny$groupwise_treatment_effects$att - attgt_ny$groupwise_treatment_effects$crit_val),
                       upper_ny = c(attgt_ny$average_treatment_effect$att + attgt_ny$average_treatment_effect$crit_val,
                                  attgt_ny$groupwise_treatment_effects$att + attgt_ny$groupwise_treatment_effects$crit_val),
                       p_value_ny = c(NA, attgt_ny$groupwise_treatment_effects$pre_treatment_p_value))

results_ny2 <- data.frame(Estimate = c("simple average", port_list$port[match(as.character(attgt_cond_ny$groupwise_treatment_effects$id), port_list$id)]),
                       ATT_cond_ny = c(attgt_cond_ny$average_treatment_effect$att,
                                   attgt_cond_ny$groupwise_treatment_effects$att),
                       lower_cond_ny = c(attgt_cond_ny$average_treatment_effect$att - attgt_cond_ny$average_treatment_effect$crit_val,
                                attgt_cond_ny$groupwise_treatment_effects$att - attgt_cond_ny$groupwise_treatment_effects$crit_val),
                       upper_cond_ny = c(attgt_cond_ny$average_treatment_effect$att + attgt_cond_ny$average_treatment_effect$crit_val,
                                  attgt_cond_ny$groupwise_treatment_effects$att + attgt_cond_ny$groupwise_treatment_effects$crit_val),
                       p_value_cond_ny = c(NA, attgt_cond_ny$groupwise_treatment_effects$pre_treatment_p_value)) 

results_ny <- merge(results_ny1, results_ny2, by.x="Estimate", by.y="Estimate", all=T,  sort=F) %>%
               mutate(sig = (ATT_ny>0 & lower_ny>0) | (ATT_ny<0 & upper_ny<0),
                      sig_cond = (ATT_cond_ny>0 & lower_cond_ny>0) | (ATT_cond_ny<0 & upper_cond_ny<0),
                      pretest = (p_value_ny < 0.05) & !is.na(p_value_ny),
                      pretest_cond = (p_value_cond_ny < 0.05) & !is.na(p_value_cond_ny),
                      across(c("ATT_ny","ATT_cond_ny","lower_ny","upper_ny","lower_cond_ny","upper_cond_ny"), ~ format(round(.,digits=0), big.mark = ",")),
                      ATT_ny = ifelse(sig==T, paste0(ATT_ny, "*"), ATT_ny),
                      ATT_ny = ifelse(pretest==T, paste0(ATT_ny, "^{x}"),ATT_ny),
                      ATT_cond_ny = ifelse(sig_cond==T, paste0(ATT_cond_ny, "*"), ATT_cond_ny),
                      ATT_cond_ny = ifelse(pretest_cond==T, paste0(ATT_cond_ny, "^{x}"),ATT_cond_ny),
                      Conf_ny = paste0("[",str_trim(lower_ny),", ",str_trim(upper_ny),"]"),
                      Conf_cond_ny = paste0("[",str_trim(lower_cond_ny),"; ",str_trim(upper_cond_ny),"]"))
results_ny <- results_ny[c("Estimate", "ATT_cond_ny", "Conf_cond_ny")]

results <- merge(results, results_ny, by.x="Estimate", by.y="Estimate", sort=F)


```

```{r include=F}
# generating a gt table
results_gt <- gt(results, rowname_col = "Estimate") |>
              tab_header(
                title = "Aggregation of group time treatment effects, treatment in period 100",
                subtitle = "Dependent variable: Total throughput with German ports in thousand TEU") |>
              cols_label(
                ATT="unconditional",
                ATT_cond="conditional",
                ATT_cond_ny="conditional",
                Conf="95% CI",
                Conf_cond="95% CI",
                Conf_cond_ny="95% CI"
              ) |>
              tab_spanner(label = "Never treated",
                          columns = c("ATT", "Conf", "ATT_cond", "Conf_cond")) |>
              tab_spanner(label = "Not yet treated",
                          columns = c("ATT_cond_ny", "Conf_cond_ny")) |>
              tab_row_group(label = "by group:",
                            rows = c(2:10)) |>
              row_group_order(groups = c(NA,"by group:")) |>
              sub_missing(columns = everything(),
                          rows = everything(),
                          missing_text = "") |>
              tab_footnote(footnote = "Stars indicate that the confidence interval does not include zero.") |>
              tab_options(stub.border.width = 0,
                          table_body.hlines.width = 0
              )

```

```{r include=TRUE, echo=FALSE}
results_gt

```

# Difference in Differences: check if Gdynia effect holds for trade with Latvia

```{r include=FALSE, warning=FALSE, message=FALSE}

loadfonts(device = "win")
set.seed(100)

#load data
load("port_data_TEU_109.RData")

port_data_TEU <- filter(port_data_TEU, aggregate==0,
                                       stat_port==1,              #filter out non-stat ports (and geographic aggregates)
                                       TEU_yrly >=20000,          #filter out small ports
                                       port_code!="DEHAM",        #exclude Hamburg due to very recent treatment
                                       sender_code=="LV") %>%      #include only containers to and from Lithuania
             mutate(s_china = ifelse(sender_code=="CN_X_HK" | sender_code=="HK", 1, 0)) %>%
             group_by(port_code, sender_code) %>%
             mutate(pair_id=cur_group_id()) %>% ungroup() %>%
             rename(container=total_TEU)


port_data_agg <- port_data_TEU %>% mutate(origin_china = ifelse(!is.na(container) & (sender_code=="CN_X_HK" | sender_code=="HK"),container,0)) %>%
                               group_by(port, port_code, period, reporter_code) %>% 
                               summarize(container = sum(container, na.rm=T),
                                         real_gdp_reporter = mean(real_gdp_reporter),
                                         origin_china = sum(origin_china),
                                         group = max(group),
                                         group_operation = max(group_operation),
                                         ownership_china = max(ownership_china),
                                         operation = max(operation),
                                         year = max(year),
                                         quarter = max(quarter)) %>% ungroup() %>%
                              group_by(port_code) %>% 
                              mutate(port_id = cur_group_id(),
                                     mean_container = mean(container, na.rm=T)) %>% ungroup()

# manipulate group variable to set treatment to September 2021 = period 100
port_data_agg <- port_data_agg %>%
  mutate(group = ifelse(group == 0, 0, 100))


#exclude Rotterdam
port_data_TEU$container[port_data_TEU$port_code=="NLRTM" & between(port_data_TEU$year,2011,2020)]<-NA
port_data_agg$container[port_data_agg$port_code=="NLRTM" & between(port_data_agg$year,2011,2020)]<-NA

port_data_agg_ny <- port_data_agg %>% filter(group!=0)

port_list <- port_data_agg %>% filter(group!=0) %>% group_by(port) %>% summarize(id = mean(port_id))

port_group_list <- port_data_agg %>% filter(group!=0) %>% group_by(group) %>%
                                     summarize(ports = toString(unique(port)))
port_group_list$ports <- paste(port_group_list$ports,paste("(", port_group_list$group, ")", sep = ""))

# Add Control variables --------------------------------------------------------

# add terminal data to the data set
#terminal_data <- read_xlsx("port_manual_berth.xlsx")

load("berth_data.RData")
terminal_data <- berth_data
  
# add channel depth and cranes from World port index
wpi_data <- read.csv("WPI_data.csv")

# control variables
control_var <- merge(wpi_data, terminal_data, by.x="port_code",by.y="port_code", all = T) %>%
               select(port_code, 
                      year, quarter, quay_length, cranes, 
                      cranes_fixed, cranes_mobile, cranes_floating, max_lift, cargo_depth)


# merge with aggregated data
port_data_agg <- merge(port_data_agg, control_var, by.x = c("port_code", "year", "quarter"), by.y = c("port_code", "year", "quarter"), all.x = T) 

#port_data_agg$ships[port_data_agg$ships==0] <- 1

port_data_agg$quay_length[port_data_agg$quay_length==0] <- 100
port_data_agg$cranes[port_data_agg$cranes==0] <- 1

### Control variables for DID ####


controls <- c("cranes", "max_lift", "cargo_depth")
var_controls <- c("quay_length")

#controls <- c("cranes", "max_lift", "cargo_depth")
#var_controls <- c("quay_length")

```



```{r include=F, message=FALSE, warning=FALSE}

#### Estimation ot ATT with never treated as control group ####
attgt <- simple_staggered_did(yname = "container",            #outcome variable
                              tname = "period",                 #time variable
                              idname = "port_id",                    #id for units (should be port_code*sender)
                              gname = "group",                    #grouping variable for first treatment. Is zero for never treated units)
                              #xformula = c("ships"),
                              #varformula = c("ships"),
                              #control_group = "not_yet_treated",
                              data = port_data_agg)


attgt_cond <- simple_staggered_did(yname = "container",            #outcome variable
                              tname = "period",                 #time variable
                              idname = "port_id",                    #id for units (should be port_code*sender)
                              gname = "group",                    #grouping variable for first treatment. Is zero for never treated units)
                              xformula = controls,
                              varformula = var_controls,
                              #control_group = "not_yet_treated",
                              data = port_data_agg)



# putting together the results

results1 <- data.frame(Estimate = c("simple average", port_list$port[match(as.character(attgt$groupwise_treatment_effects$id), port_list$id)]),
                       ATT = c(attgt$average_treatment_effect$att,
                                   attgt$groupwise_treatment_effects$att),
                       lower = c(attgt$average_treatment_effect$att - attgt$average_treatment_effect$crit_val,
                                attgt$groupwise_treatment_effects$att - attgt$groupwise_treatment_effects$crit_val),
                       upper = c(attgt$average_treatment_effect$att + attgt$average_treatment_effect$crit_val,
                                  attgt$groupwise_treatment_effects$att + attgt$groupwise_treatment_effects$crit_val),
                       p_value = c(NA, attgt$groupwise_treatment_effects$pre_treatment_p_value))

results2 <- data.frame(Estimate = c("simple average", port_list$port[match(as.character(attgt_cond$groupwise_treatment_effects$id), port_list$id)]),
                       ATT_cond = c(attgt_cond$average_treatment_effect$att,
                                   attgt_cond$groupwise_treatment_effects$att),
                       lower_cond = c(attgt_cond$average_treatment_effect$att - attgt_cond$average_treatment_effect$crit_val,
                                attgt_cond$groupwise_treatment_effects$att - attgt_cond$groupwise_treatment_effects$crit_val),
                       upper_cond = c(attgt_cond$average_treatment_effect$att + attgt_cond$average_treatment_effect$crit_val,
                                  attgt_cond$groupwise_treatment_effects$att + attgt_cond$groupwise_treatment_effects$crit_val),
                        p_value_cond = c(NA, attgt_cond$groupwise_treatment_effects$pre_treatment_p_value))

results <- merge(results1, results2, by.x="Estimate", by.y="Estimate", all=T,  sort=F) %>%
               mutate(sig = (ATT>0 & lower>0) | (ATT<0 & upper<0),
                      sig_cond = (ATT_cond>0 & lower_cond>0) | (ATT_cond<0 & upper_cond<0),
                      pretest = (p_value < 0.05) & !is.na(p_value),
                      pretest_cond = (p_value_cond < 0.05) & !is.na(p_value_cond),
                      across(c("ATT","ATT_cond","lower","upper","lower_cond","upper_cond"), ~ format(round(.,digits=0), big.mark = ",")),

                      ATT = ifelse(sig==T, paste0(ATT, "*"), ATT),
                      ATT = ifelse(pretest==T, paste0(ATT, "^{x}"),ATT),
                      ATT_cond = ifelse(sig_cond==T, paste0(ATT_cond, "*"), ATT_cond),
                      ATT_cond = ifelse(pretest_cond==T, paste0(ATT_cond, "^{x}"),ATT_cond),
                      Conf = paste0("[",str_trim(lower),", ",str_trim(upper),"]"),
                      Conf_cond = paste0("[",str_trim(lower_cond),"; ",str_trim(upper_cond),"]"))
results <- results[c("Estimate", "ATT", "Conf", "ATT_cond", "Conf_cond")]



#### Estimate ATT with not-yet-treated control group ####
attgt_ny <- simple_staggered_did(yname = "container",            #outcome variable
                              tname = "period",                 #time variable
                              idname = "port_id",                    #id for units (should be port_code*sender)
                              gname = "group",                    #grouping variable for first treatment. Is zero for never treated units)
                              #xformula = c("ships"),
                              #varformula = c("ships"),
                              control_group = "not_yet_treated",
                              data = port_data_agg)


attgt_cond_ny <- simple_staggered_did(yname = "container",            #outcome variable
                              tname = "period",                 #time variable
                              idname = "port_id",                    #id for units (should be port_code*sender)
                              gname = "group",                    #grouping variable for first treatment. Is zero for never treated units)
                              xformula = controls,
                              varformula = var_controls,
                              control_group = "not_yet_treated",
                              data = port_data_agg)



# putting together the results

results_ny1 <- data.frame(Estimate = c("simple average", port_list$port[match(as.character(attgt_ny$groupwise_treatment_effects$id), port_list$id)]),
                       ATT_ny = c(attgt_ny$average_treatment_effect$att,
                                   attgt_ny$groupwise_treatment_effects$att),
                       lower_ny = c(attgt_ny$average_treatment_effect$att - attgt_ny$average_treatment_effect$crit_val,
                                attgt_ny$groupwise_treatment_effects$att - attgt_ny$groupwise_treatment_effects$crit_val),
                       upper_ny = c(attgt_ny$average_treatment_effect$att + attgt_ny$average_treatment_effect$crit_val,
                                  attgt_ny$groupwise_treatment_effects$att + attgt_ny$groupwise_treatment_effects$crit_val),
                       p_value_ny = c(NA, attgt_ny$groupwise_treatment_effects$pre_treatment_p_value))

results_ny2 <- data.frame(Estimate = c("simple average", port_list$port[match(as.character(attgt_cond_ny$groupwise_treatment_effects$id), port_list$id)]),
                       ATT_cond_ny = c(attgt_cond_ny$average_treatment_effect$att,
                                   attgt_cond_ny$groupwise_treatment_effects$att),
                       lower_cond_ny = c(attgt_cond_ny$average_treatment_effect$att - attgt_cond_ny$average_treatment_effect$crit_val,
                                attgt_cond_ny$groupwise_treatment_effects$att - attgt_cond_ny$groupwise_treatment_effects$crit_val),
                       upper_cond_ny = c(attgt_cond_ny$average_treatment_effect$att + attgt_cond_ny$average_treatment_effect$crit_val,
                                  attgt_cond_ny$groupwise_treatment_effects$att + attgt_cond_ny$groupwise_treatment_effects$crit_val),
                       p_value_cond_ny = c(NA, attgt_cond_ny$groupwise_treatment_effects$pre_treatment_p_value)) 

results_ny <- merge(results_ny1, results_ny2, by.x="Estimate", by.y="Estimate", all=T,  sort=F) %>%
               mutate(sig = (ATT_ny>0 & lower_ny>0) | (ATT_ny<0 & upper_ny<0),
                      sig_cond = (ATT_cond_ny>0 & lower_cond_ny>0) | (ATT_cond_ny<0 & upper_cond_ny<0),
                      pretest = (p_value_ny < 0.05) & !is.na(p_value_ny),
                      pretest_cond = (p_value_cond_ny < 0.05) & !is.na(p_value_cond_ny),
                      across(c("ATT_ny","ATT_cond_ny","lower_ny","upper_ny","lower_cond_ny","upper_cond_ny"), ~ format(round(.,digits=0), big.mark = ",")),
                      ATT_ny = ifelse(sig==T, paste0(ATT_ny, "*"), ATT_ny),
                      ATT_ny = ifelse(pretest==T, paste0(ATT_ny, "^{x}"),ATT_ny),
                      ATT_cond_ny = ifelse(sig_cond==T, paste0(ATT_cond_ny, "*"), ATT_cond_ny),
                      ATT_cond_ny = ifelse(pretest_cond==T, paste0(ATT_cond_ny, "^{x}"),ATT_cond_ny),
                      Conf_ny = paste0("[",str_trim(lower_ny),", ",str_trim(upper_ny),"]"),
                      Conf_cond_ny = paste0("[",str_trim(lower_cond_ny),"; ",str_trim(upper_cond_ny),"]"))
results_ny <- results_ny[c("Estimate", "ATT_cond_ny", "Conf_cond_ny")]

results <- merge(results, results_ny, by.x="Estimate", by.y="Estimate", sort=F)


```

```{r include=F}
# generating a gt table
results_gt <- gt(results, rowname_col = "Estimate") |>
              tab_header(
                title = "Aggregation of group time treatment effects, treatment in period 100",
                subtitle = "Dependent variable: Total throughput with Latvian ports in thousand TEU") |>
              cols_label(
                ATT="unconditional",
                ATT_cond="conditional",
                ATT_cond_ny="conditional",
                Conf="95% CI",
                Conf_cond="95% CI",
                Conf_cond_ny="95% CI"
              ) |>
              tab_spanner(label = "Never treated",
                          columns = c("ATT", "Conf", "ATT_cond", "Conf_cond")) |>
              tab_spanner(label = "Not yet treated",
                          columns = c("ATT_cond_ny", "Conf_cond_ny")) |>
              tab_row_group(label = "by group:",
                            rows = c(2:10)) |>
              row_group_order(groups = c(NA,"by group:")) |>
              sub_missing(columns = everything(),
                          rows = everything(),
                          missing_text = "") |>
              tab_footnote(footnote = "Stars indicate that the confidence interval does not include zero.") |>
              tab_options(stub.border.width = 0,
                          table_body.hlines.width = 0
              )

```

```{r include=TRUE, echo=FALSE}
results_gt

```

The negative Gdynia treatment effect at period 100 (September 2021) can not be found for trade with German ports (western neighbor), and also with Latvian ports (eastern neighbor). Maybe there is indeed a policy tool effect that is local. The results so far suggest that Gdynia experienced a decline in trade with Lithuania after treatment that was not trivial for its overall performance and that was not seen with trade with other important partners: Germany and Latvia.


# Difference in Differences: cross-port variation in trade with ports in Lithuania - covid treatment in period 94

```{r include=FALSE, warning=FALSE, message=FALSE}

loadfonts(device = "win")
set.seed(100)

#load data
load("port_data_TEU_109.RData")

port_data_TEU <- filter(port_data_TEU, aggregate==0,
                                       stat_port==1,              #filter out non-stat ports (and geographic aggregates)
                                       TEU_yrly >=20000,          #filter out small ports
                                       port_code!="DEHAM",        #exclude Hamburg due to very recent treatment
                                       sender_code=="LT") %>%      #include only containers to and from Lithuania
             mutate(s_china = ifelse(sender_code=="CN_X_HK" | sender_code=="HK", 1, 0)) %>%
             group_by(port_code, sender_code) %>%
             mutate(pair_id=cur_group_id()) %>% ungroup() %>%
             rename(container=total_TEU)


port_data_agg <- port_data_TEU %>% mutate(origin_china = ifelse(!is.na(container) & (sender_code=="CN_X_HK" | sender_code=="HK"),container,0)) %>%
                               group_by(port, port_code, period, reporter_code) %>% 
                               summarize(container = sum(container, na.rm=T),
                                         real_gdp_reporter = mean(real_gdp_reporter),
                                         origin_china = sum(origin_china),
                                         group = max(group),
                                         group_operation = max(group_operation),
                                         ownership_china = max(ownership_china),
                                         operation = max(operation),
                                         year = max(year),
                                         quarter = max(quarter)) %>% ungroup() %>%
                              group_by(port_code) %>% 
                              mutate(port_id = cur_group_id(),
                                     mean_container = mean(container, na.rm=T)) %>% ungroup()

# manipulate group variable to set treatment to September 2021 = period 100
port_data_agg <- port_data_agg %>%
  mutate(group = ifelse(group == 0, 0, 94))


#exclude Rotterdam
port_data_TEU$container[port_data_TEU$port_code=="NLRTM" & between(port_data_TEU$year,2011,2020)]<-NA
port_data_agg$container[port_data_agg$port_code=="NLRTM" & between(port_data_agg$year,2011,2020)]<-NA

port_data_agg_ny <- port_data_agg %>% filter(group!=0)

port_list <- port_data_agg %>% filter(group!=0) %>% group_by(port) %>% summarize(id = mean(port_id))

port_group_list <- port_data_agg %>% filter(group!=0) %>% group_by(group) %>%
                                     summarize(ports = toString(unique(port)))
port_group_list$ports <- paste(port_group_list$ports,paste("(", port_group_list$group, ")", sep = ""))

# Add Control variables --------------------------------------------------------

# add terminal data to the data set
#terminal_data <- read_xlsx("port_manual_berth.xlsx")

load("berth_data.RData")
terminal_data <- berth_data
  
# add channel depth and cranes from World port index
wpi_data <- read.csv("WPI_data.csv")

# control variables
control_var <- merge(wpi_data, terminal_data, by.x="port_code",by.y="port_code", all = T) %>%
               select(port_code, 
                      year, quarter, quay_length, cranes, 
                      cranes_fixed, cranes_mobile, cranes_floating, max_lift, cargo_depth)


# merge with aggregated data
port_data_agg <- merge(port_data_agg, control_var, by.x = c("port_code", "year", "quarter"), by.y = c("port_code", "year", "quarter"), all.x = T) 

#port_data_agg$ships[port_data_agg$ships==0] <- 1

port_data_agg$quay_length[port_data_agg$quay_length==0] <- 100
port_data_agg$cranes[port_data_agg$cranes==0] <- 1

### Control variables for DID ####


controls <- c("cranes", "max_lift", "cargo_depth")
var_controls <- c("quay_length")

#controls <- c("cranes", "max_lift", "cargo_depth")
#var_controls <- c("quay_length")

```



```{r include=F, message=FALSE, warning=FALSE}

#### Estimation ot ATT with never treated as control group ####
attgt <- simple_staggered_did(yname = "container",            #outcome variable
                              tname = "period",                 #time variable
                              idname = "port_id",                    #id for units (should be port_code*sender)
                              gname = "group",                    #grouping variable for first treatment. Is zero for never treated units)
                              #xformula = c("ships"),
                              #varformula = c("ships"),
                              #control_group = "not_yet_treated",
                              data = port_data_agg)


attgt_cond <- simple_staggered_did(yname = "container",            #outcome variable
                              tname = "period",                 #time variable
                              idname = "port_id",                    #id for units (should be port_code*sender)
                              gname = "group",                    #grouping variable for first treatment. Is zero for never treated units)
                              xformula = controls,
                              varformula = var_controls,
                              #control_group = "not_yet_treated",
                              data = port_data_agg)



# putting together the results

results1 <- data.frame(Estimate = c("simple average", port_list$port[match(as.character(attgt$groupwise_treatment_effects$id), port_list$id)]),
                       ATT = c(attgt$average_treatment_effect$att,
                                   attgt$groupwise_treatment_effects$att),
                       lower = c(attgt$average_treatment_effect$att - attgt$average_treatment_effect$crit_val,
                                attgt$groupwise_treatment_effects$att - attgt$groupwise_treatment_effects$crit_val),
                       upper = c(attgt$average_treatment_effect$att + attgt$average_treatment_effect$crit_val,
                                  attgt$groupwise_treatment_effects$att + attgt$groupwise_treatment_effects$crit_val),
                       p_value = c(NA, attgt$groupwise_treatment_effects$pre_treatment_p_value))

results2 <- data.frame(Estimate = c("simple average", port_list$port[match(as.character(attgt_cond$groupwise_treatment_effects$id), port_list$id)]),
                       ATT_cond = c(attgt_cond$average_treatment_effect$att,
                                   attgt_cond$groupwise_treatment_effects$att),
                       lower_cond = c(attgt_cond$average_treatment_effect$att - attgt_cond$average_treatment_effect$crit_val,
                                attgt_cond$groupwise_treatment_effects$att - attgt_cond$groupwise_treatment_effects$crit_val),
                       upper_cond = c(attgt_cond$average_treatment_effect$att + attgt_cond$average_treatment_effect$crit_val,
                                  attgt_cond$groupwise_treatment_effects$att + attgt_cond$groupwise_treatment_effects$crit_val),
                        p_value_cond = c(NA, attgt_cond$groupwise_treatment_effects$pre_treatment_p_value))

results <- merge(results1, results2, by.x="Estimate", by.y="Estimate", all=T,  sort=F) %>%
               mutate(sig = (ATT>0 & lower>0) | (ATT<0 & upper<0),
                      sig_cond = (ATT_cond>0 & lower_cond>0) | (ATT_cond<0 & upper_cond<0),
                      pretest = (p_value < 0.05) & !is.na(p_value),
                      pretest_cond = (p_value_cond < 0.05) & !is.na(p_value_cond),
                      across(c("ATT","ATT_cond","lower","upper","lower_cond","upper_cond"), ~ format(round(.,digits=0), big.mark = ",")),

                      ATT = ifelse(sig==T, paste0(ATT, "*"), ATT),
                      ATT = ifelse(pretest==T, paste0(ATT, "^{x}"),ATT),
                      ATT_cond = ifelse(sig_cond==T, paste0(ATT_cond, "*"), ATT_cond),
                      ATT_cond = ifelse(pretest_cond==T, paste0(ATT_cond, "^{x}"),ATT_cond),
                      Conf = paste0("[",str_trim(lower),", ",str_trim(upper),"]"),
                      Conf_cond = paste0("[",str_trim(lower_cond),"; ",str_trim(upper_cond),"]"))
results <- results[c("Estimate", "ATT", "Conf", "ATT_cond", "Conf_cond")]



#### Estimate ATT with not-yet-treated control group ####
attgt_ny <- simple_staggered_did(yname = "container",            #outcome variable
                              tname = "period",                 #time variable
                              idname = "port_id",                    #id for units (should be port_code*sender)
                              gname = "group",                    #grouping variable for first treatment. Is zero for never treated units)
                              #xformula = c("ships"),
                              #varformula = c("ships"),
                              control_group = "not_yet_treated",
                              data = port_data_agg)


attgt_cond_ny <- simple_staggered_did(yname = "container",            #outcome variable
                              tname = "period",                 #time variable
                              idname = "port_id",                    #id for units (should be port_code*sender)
                              gname = "group",                    #grouping variable for first treatment. Is zero for never treated units)
                              xformula = controls,
                              varformula = var_controls,
                              control_group = "not_yet_treated",
                              data = port_data_agg)



# putting together the results

results_ny1 <- data.frame(Estimate = c("simple average", port_list$port[match(as.character(attgt_ny$groupwise_treatment_effects$id), port_list$id)]),
                       ATT_ny = c(attgt_ny$average_treatment_effect$att,
                                   attgt_ny$groupwise_treatment_effects$att),
                       lower_ny = c(attgt_ny$average_treatment_effect$att - attgt_ny$average_treatment_effect$crit_val,
                                attgt_ny$groupwise_treatment_effects$att - attgt_ny$groupwise_treatment_effects$crit_val),
                       upper_ny = c(attgt_ny$average_treatment_effect$att + attgt_ny$average_treatment_effect$crit_val,
                                  attgt_ny$groupwise_treatment_effects$att + attgt_ny$groupwise_treatment_effects$crit_val),
                       p_value_ny = c(NA, attgt_ny$groupwise_treatment_effects$pre_treatment_p_value))

results_ny2 <- data.frame(Estimate = c("simple average", port_list$port[match(as.character(attgt_cond_ny$groupwise_treatment_effects$id), port_list$id)]),
                       ATT_cond_ny = c(attgt_cond_ny$average_treatment_effect$att,
                                   attgt_cond_ny$groupwise_treatment_effects$att),
                       lower_cond_ny = c(attgt_cond_ny$average_treatment_effect$att - attgt_cond_ny$average_treatment_effect$crit_val,
                                attgt_cond_ny$groupwise_treatment_effects$att - attgt_cond_ny$groupwise_treatment_effects$crit_val),
                       upper_cond_ny = c(attgt_cond_ny$average_treatment_effect$att + attgt_cond_ny$average_treatment_effect$crit_val,
                                  attgt_cond_ny$groupwise_treatment_effects$att + attgt_cond_ny$groupwise_treatment_effects$crit_val),
                       p_value_cond_ny = c(NA, attgt_cond_ny$groupwise_treatment_effects$pre_treatment_p_value)) 

results_ny <- merge(results_ny1, results_ny2, by.x="Estimate", by.y="Estimate", all=T,  sort=F) %>%
               mutate(sig = (ATT_ny>0 & lower_ny>0) | (ATT_ny<0 & upper_ny<0),
                      sig_cond = (ATT_cond_ny>0 & lower_cond_ny>0) | (ATT_cond_ny<0 & upper_cond_ny<0),
                      pretest = (p_value_ny < 0.05) & !is.na(p_value_ny),
                      pretest_cond = (p_value_cond_ny < 0.05) & !is.na(p_value_cond_ny),
                      across(c("ATT_ny","ATT_cond_ny","lower_ny","upper_ny","lower_cond_ny","upper_cond_ny"), ~ format(round(.,digits=0), big.mark = ",")),
                      ATT_ny = ifelse(sig==T, paste0(ATT_ny, "*"), ATT_ny),
                      ATT_ny = ifelse(pretest==T, paste0(ATT_ny, "^{x}"),ATT_ny),
                      ATT_cond_ny = ifelse(sig_cond==T, paste0(ATT_cond_ny, "*"), ATT_cond_ny),
                      ATT_cond_ny = ifelse(pretest_cond==T, paste0(ATT_cond_ny, "^{x}"),ATT_cond_ny),
                      Conf_ny = paste0("[",str_trim(lower_ny),", ",str_trim(upper_ny),"]"),
                      Conf_cond_ny = paste0("[",str_trim(lower_cond_ny),"; ",str_trim(upper_cond_ny),"]"))
results_ny <- results_ny[c("Estimate", "ATT_cond_ny", "Conf_cond_ny")]

results <- merge(results, results_ny, by.x="Estimate", by.y="Estimate", sort=F)


```

```{r include=F}
# generating a gt table
results_gt <- gt(results, rowname_col = "Estimate") |>
              tab_header(
                title = "Aggregation of group time treatment effects, placebo treatment in period 94 (covid treatment 2020 March)",
                subtitle = "Dependent variable: Total throughput with Lithuania in thousand TEU") |>
              cols_label(
                ATT="unconditional",
                ATT_cond="conditional",
                ATT_cond_ny="conditional",
                Conf="95% CI",
                Conf_cond="95% CI",
                Conf_cond_ny="95% CI"
              ) |>
              tab_spanner(label = "Never treated",
                          columns = c("ATT", "Conf", "ATT_cond", "Conf_cond")) |>
              tab_spanner(label = "Not yet treated",
                          columns = c("ATT_cond_ny", "Conf_cond_ny")) |>
              tab_row_group(label = "by group:",
                            rows = c(2:10)) |>
              row_group_order(groups = c(NA,"by group:")) |>
              sub_missing(columns = everything(),
                          rows = everything(),
                          missing_text = "") |>
              tab_footnote(footnote = "Stars indicate that the confidence interval does not include zero.") |>
              tab_options(stub.border.width = 0,
                          table_body.hlines.width = 0
              )

```

```{r include=TRUE, echo=FALSE}
results_gt
```

Was the negative treatment effect with 100 rather a covid-19 effect kicking in at period 94 than a policy effect? Maybe partly but not all if we compare the size of the effects and assume that the covid effect was an instant effect that faded out over the periods rather than accumulated and increased. I would find the previous development of the covid effect (jump and then fade out) more intuitive.

# Difference in Differences: cross-port variation in trade with ports in Lithuania: Placebo treatment in period 90

```{r include=FALSE, warning=FALSE, message=FALSE}

loadfonts(device = "win")
set.seed(100)

#load data
load("port_data_TEU_109.RData")

port_data_TEU <- filter(port_data_TEU, aggregate==0,
                                       stat_port==1,              #filter out non-stat ports (and geographic aggregates)
                                       TEU_yrly >=20000,          #filter out small ports
                                       port_code!="DEHAM",        #exclude Hamburg due to very recent treatment
                                       sender_code=="LT") %>%      #include only containers to and from Lithuania
             mutate(s_china = ifelse(sender_code=="CN_X_HK" | sender_code=="HK", 1, 0)) %>%
             group_by(port_code, sender_code) %>%
             mutate(pair_id=cur_group_id()) %>% ungroup() %>%
             rename(container=total_TEU)


port_data_agg <- port_data_TEU %>% mutate(origin_china = ifelse(!is.na(container) & (sender_code=="CN_X_HK" | sender_code=="HK"),container,0)) %>%
                               group_by(port, port_code, period, reporter_code) %>% 
                               summarize(container = sum(container, na.rm=T),
                                         real_gdp_reporter = mean(real_gdp_reporter),
                                         origin_china = sum(origin_china),
                                         group = max(group),
                                         group_operation = max(group_operation),
                                         ownership_china = max(ownership_china),
                                         operation = max(operation),
                                         year = max(year),
                                         quarter = max(quarter)) %>% ungroup() %>%
                              group_by(port_code) %>% 
                              mutate(port_id = cur_group_id(),
                                     mean_container = mean(container, na.rm=T)) %>% ungroup()

# manipulate group variable to set treatment to September 2021 = period 100
port_data_agg <- port_data_agg %>%
  mutate(group = ifelse(group == 0, 0, 90))


#exclude Rotterdam
port_data_TEU$container[port_data_TEU$port_code=="NLRTM" & between(port_data_TEU$year,2011,2020)]<-NA
port_data_agg$container[port_data_agg$port_code=="NLRTM" & between(port_data_agg$year,2011,2020)]<-NA

port_data_agg_ny <- port_data_agg %>% filter(group!=0)

port_list <- port_data_agg %>% filter(group!=0) %>% group_by(port) %>% summarize(id = mean(port_id))

port_group_list <- port_data_agg %>% filter(group!=0) %>% group_by(group) %>%
                                     summarize(ports = toString(unique(port)))
port_group_list$ports <- paste(port_group_list$ports,paste("(", port_group_list$group, ")", sep = ""))

# Add Control variables --------------------------------------------------------

# add terminal data to the data set
#terminal_data <- read_xlsx("port_manual_berth.xlsx")

load("berth_data.RData")
terminal_data <- berth_data
  
# add channel depth and cranes from World port index
wpi_data <- read.csv("WPI_data.csv")

# control variables
control_var <- merge(wpi_data, terminal_data, by.x="port_code",by.y="port_code", all = T) %>%
               select(port_code, 
                      year, quarter, quay_length, cranes, 
                      cranes_fixed, cranes_mobile, cranes_floating, max_lift, cargo_depth)


# merge with aggregated data
port_data_agg <- merge(port_data_agg, control_var, by.x = c("port_code", "year", "quarter"), by.y = c("port_code", "year", "quarter"), all.x = T) 

#port_data_agg$ships[port_data_agg$ships==0] <- 1

port_data_agg$quay_length[port_data_agg$quay_length==0] <- 100
port_data_agg$cranes[port_data_agg$cranes==0] <- 1

### Control variables for DID ####


controls <- c("cranes", "max_lift", "cargo_depth")
var_controls <- c("quay_length")

#controls <- c("cranes", "max_lift", "cargo_depth")
#var_controls <- c("quay_length")

```



```{r include=F, message=FALSE, warning=FALSE}

#### Estimation ot ATT with never treated as control group ####
attgt <- simple_staggered_did(yname = "container",            #outcome variable
                              tname = "period",                 #time variable
                              idname = "port_id",                    #id for units (should be port_code*sender)
                              gname = "group",                    #grouping variable for first treatment. Is zero for never treated units)
                              #xformula = c("ships"),
                              #varformula = c("ships"),
                              #control_group = "not_yet_treated",
                              data = port_data_agg)


attgt_cond <- simple_staggered_did(yname = "container",            #outcome variable
                              tname = "period",                 #time variable
                              idname = "port_id",                    #id for units (should be port_code*sender)
                              gname = "group",                    #grouping variable for first treatment. Is zero for never treated units)
                              xformula = controls,
                              varformula = var_controls,
                              #control_group = "not_yet_treated",
                              data = port_data_agg)



# putting together the results

results1 <- data.frame(Estimate = c("simple average", port_list$port[match(as.character(attgt$groupwise_treatment_effects$id), port_list$id)]),
                       ATT = c(attgt$average_treatment_effect$att,
                                   attgt$groupwise_treatment_effects$att),
                       lower = c(attgt$average_treatment_effect$att - attgt$average_treatment_effect$crit_val,
                                attgt$groupwise_treatment_effects$att - attgt$groupwise_treatment_effects$crit_val),
                       upper = c(attgt$average_treatment_effect$att + attgt$average_treatment_effect$crit_val,
                                  attgt$groupwise_treatment_effects$att + attgt$groupwise_treatment_effects$crit_val),
                       p_value = c(NA, attgt$groupwise_treatment_effects$pre_treatment_p_value))

results2 <- data.frame(Estimate = c("simple average", port_list$port[match(as.character(attgt_cond$groupwise_treatment_effects$id), port_list$id)]),
                       ATT_cond = c(attgt_cond$average_treatment_effect$att,
                                   attgt_cond$groupwise_treatment_effects$att),
                       lower_cond = c(attgt_cond$average_treatment_effect$att - attgt_cond$average_treatment_effect$crit_val,
                                attgt_cond$groupwise_treatment_effects$att - attgt_cond$groupwise_treatment_effects$crit_val),
                       upper_cond = c(attgt_cond$average_treatment_effect$att + attgt_cond$average_treatment_effect$crit_val,
                                  attgt_cond$groupwise_treatment_effects$att + attgt_cond$groupwise_treatment_effects$crit_val),
                        p_value_cond = c(NA, attgt_cond$groupwise_treatment_effects$pre_treatment_p_value))

results <- merge(results1, results2, by.x="Estimate", by.y="Estimate", all=T,  sort=F) %>%
               mutate(sig = (ATT>0 & lower>0) | (ATT<0 & upper<0),
                      sig_cond = (ATT_cond>0 & lower_cond>0) | (ATT_cond<0 & upper_cond<0),
                      pretest = (p_value < 0.05) & !is.na(p_value),
                      pretest_cond = (p_value_cond < 0.05) & !is.na(p_value_cond),
                      across(c("ATT","ATT_cond","lower","upper","lower_cond","upper_cond"), ~ format(round(.,digits=0), big.mark = ",")),

                      ATT = ifelse(sig==T, paste0(ATT, "*"), ATT),
                      ATT = ifelse(pretest==T, paste0(ATT, "^{x}"),ATT),
                      ATT_cond = ifelse(sig_cond==T, paste0(ATT_cond, "*"), ATT_cond),
                      ATT_cond = ifelse(pretest_cond==T, paste0(ATT_cond, "^{x}"),ATT_cond),
                      Conf = paste0("[",str_trim(lower),", ",str_trim(upper),"]"),
                      Conf_cond = paste0("[",str_trim(lower_cond),"; ",str_trim(upper_cond),"]"))
results <- results[c("Estimate", "ATT", "Conf", "ATT_cond", "Conf_cond")]



#### Estimate ATT with not-yet-treated control group ####
attgt_ny <- simple_staggered_did(yname = "container",            #outcome variable
                              tname = "period",                 #time variable
                              idname = "port_id",                    #id for units (should be port_code*sender)
                              gname = "group",                    #grouping variable for first treatment. Is zero for never treated units)
                              #xformula = c("ships"),
                              #varformula = c("ships"),
                              control_group = "not_yet_treated",
                              data = port_data_agg)


attgt_cond_ny <- simple_staggered_did(yname = "container",            #outcome variable
                              tname = "period",                 #time variable
                              idname = "port_id",                    #id for units (should be port_code*sender)
                              gname = "group",                    #grouping variable for first treatment. Is zero for never treated units)
                              xformula = controls,
                              varformula = var_controls,
                              control_group = "not_yet_treated",
                              data = port_data_agg)



# putting together the results

results_ny1 <- data.frame(Estimate = c("simple average", port_list$port[match(as.character(attgt_ny$groupwise_treatment_effects$id), port_list$id)]),
                       ATT_ny = c(attgt_ny$average_treatment_effect$att,
                                   attgt_ny$groupwise_treatment_effects$att),
                       lower_ny = c(attgt_ny$average_treatment_effect$att - attgt_ny$average_treatment_effect$crit_val,
                                attgt_ny$groupwise_treatment_effects$att - attgt_ny$groupwise_treatment_effects$crit_val),
                       upper_ny = c(attgt_ny$average_treatment_effect$att + attgt_ny$average_treatment_effect$crit_val,
                                  attgt_ny$groupwise_treatment_effects$att + attgt_ny$groupwise_treatment_effects$crit_val),
                       p_value_ny = c(NA, attgt_ny$groupwise_treatment_effects$pre_treatment_p_value))

results_ny2 <- data.frame(Estimate = c("simple average", port_list$port[match(as.character(attgt_cond_ny$groupwise_treatment_effects$id), port_list$id)]),
                       ATT_cond_ny = c(attgt_cond_ny$average_treatment_effect$att,
                                   attgt_cond_ny$groupwise_treatment_effects$att),
                       lower_cond_ny = c(attgt_cond_ny$average_treatment_effect$att - attgt_cond_ny$average_treatment_effect$crit_val,
                                attgt_cond_ny$groupwise_treatment_effects$att - attgt_cond_ny$groupwise_treatment_effects$crit_val),
                       upper_cond_ny = c(attgt_cond_ny$average_treatment_effect$att + attgt_cond_ny$average_treatment_effect$crit_val,
                                  attgt_cond_ny$groupwise_treatment_effects$att + attgt_cond_ny$groupwise_treatment_effects$crit_val),
                       p_value_cond_ny = c(NA, attgt_cond_ny$groupwise_treatment_effects$pre_treatment_p_value)) 

results_ny <- merge(results_ny1, results_ny2, by.x="Estimate", by.y="Estimate", all=T,  sort=F) %>%
               mutate(sig = (ATT_ny>0 & lower_ny>0) | (ATT_ny<0 & upper_ny<0),
                      sig_cond = (ATT_cond_ny>0 & lower_cond_ny>0) | (ATT_cond_ny<0 & upper_cond_ny<0),
                      pretest = (p_value_ny < 0.05) & !is.na(p_value_ny),
                      pretest_cond = (p_value_cond_ny < 0.05) & !is.na(p_value_cond_ny),
                      across(c("ATT_ny","ATT_cond_ny","lower_ny","upper_ny","lower_cond_ny","upper_cond_ny"), ~ format(round(.,digits=0), big.mark = ",")),
                      ATT_ny = ifelse(sig==T, paste0(ATT_ny, "*"), ATT_ny),
                      ATT_ny = ifelse(pretest==T, paste0(ATT_ny, "^{x}"),ATT_ny),
                      ATT_cond_ny = ifelse(sig_cond==T, paste0(ATT_cond_ny, "*"), ATT_cond_ny),
                      ATT_cond_ny = ifelse(pretest_cond==T, paste0(ATT_cond_ny, "^{x}"),ATT_cond_ny),
                      Conf_ny = paste0("[",str_trim(lower_ny),", ",str_trim(upper_ny),"]"),
                      Conf_cond_ny = paste0("[",str_trim(lower_cond_ny),"; ",str_trim(upper_cond_ny),"]"))
results_ny <- results_ny[c("Estimate", "ATT_cond_ny", "Conf_cond_ny")]

results <- merge(results, results_ny, by.x="Estimate", by.y="Estimate", sort=F)


```

```{r include=F}
# generating a gt table
results_gt <- gt(results, rowname_col = "Estimate") |>
              tab_header(
                title = "Aggregation of group time treatment effects, placebo treatment in period 90",
                subtitle = "Dependent variable: Total throughput with Lithuania in thousand TEU") |>
              cols_label(
                ATT="unconditional",
                ATT_cond="conditional",
                ATT_cond_ny="conditional",
                Conf="95% CI",
                Conf_cond="95% CI",
                Conf_cond_ny="95% CI"
              ) |>
              tab_spanner(label = "Never treated",
                          columns = c("ATT", "Conf", "ATT_cond", "Conf_cond")) |>
              tab_spanner(label = "Not yet treated",
                          columns = c("ATT_cond_ny", "Conf_cond_ny")) |>
              tab_row_group(label = "by group:",
                            rows = c(2:10)) |>
              row_group_order(groups = c(NA,"by group:")) |>
              sub_missing(columns = everything(),
                          rows = everything(),
                          missing_text = "") |>
              tab_footnote(footnote = "Stars indicate that the confidence interval does not include zero.") |>
              tab_options(stub.border.width = 0,
                          table_body.hlines.width = 0
              )

```

```{r include=TRUE, echo=FALSE}
results_gt
```

# Difference in Differences: cross-port variation in trade with ports in Lithuania: Placebo treatment in period 80

```{r include=FALSE, warning=FALSE, message=FALSE}

loadfonts(device = "win")
set.seed(100)

#load data
load("port_data_TEU_109.RData")

port_data_TEU <- filter(port_data_TEU, aggregate==0,
                                       stat_port==1,              #filter out non-stat ports (and geographic aggregates)
                                       TEU_yrly >=20000,          #filter out small ports
                                       port_code!="DEHAM",        #exclude Hamburg due to very recent treatment
                                       sender_code=="LT") %>%      #include only containers to and from Lithuania
             mutate(s_china = ifelse(sender_code=="CN_X_HK" | sender_code=="HK", 1, 0)) %>%
             group_by(port_code, sender_code) %>%
             mutate(pair_id=cur_group_id()) %>% ungroup() %>%
             rename(container=total_TEU)


port_data_agg <- port_data_TEU %>% mutate(origin_china = ifelse(!is.na(container) & (sender_code=="CN_X_HK" | sender_code=="HK"),container,0)) %>%
                               group_by(port, port_code, period, reporter_code) %>% 
                               summarize(container = sum(container, na.rm=T),
                                         real_gdp_reporter = mean(real_gdp_reporter),
                                         origin_china = sum(origin_china),
                                         group = max(group),
                                         group_operation = max(group_operation),
                                         ownership_china = max(ownership_china),
                                         operation = max(operation),
                                         year = max(year),
                                         quarter = max(quarter)) %>% ungroup() %>%
                              group_by(port_code) %>% 
                              mutate(port_id = cur_group_id(),
                                     mean_container = mean(container, na.rm=T)) %>% ungroup()

# manipulate group variable to set treatment to September 2021 = period 100
port_data_agg <- port_data_agg %>%
  mutate(group = ifelse(group == 0, 0, 90))


#exclude Rotterdam
port_data_TEU$container[port_data_TEU$port_code=="NLRTM" & between(port_data_TEU$year,2011,2020)]<-NA
port_data_agg$container[port_data_agg$port_code=="NLRTM" & between(port_data_agg$year,2011,2020)]<-NA

port_data_agg_ny <- port_data_agg %>% filter(group!=0)

port_list <- port_data_agg %>% filter(group!=0) %>% group_by(port) %>% summarize(id = mean(port_id))

port_group_list <- port_data_agg %>% filter(group!=0) %>% group_by(group) %>%
                                     summarize(ports = toString(unique(port)))
port_group_list$ports <- paste(port_group_list$ports,paste("(", port_group_list$group, ")", sep = ""))

# Add Control variables --------------------------------------------------------

# add terminal data to the data set
#terminal_data <- read_xlsx("port_manual_berth.xlsx")

load("berth_data.RData")
terminal_data <- berth_data
  
# add channel depth and cranes from World port index
wpi_data <- read.csv("WPI_data.csv")

# control variables
control_var <- merge(wpi_data, terminal_data, by.x="port_code",by.y="port_code", all = T) %>%
               select(port_code, 
                      year, quarter, quay_length, cranes, 
                      cranes_fixed, cranes_mobile, cranes_floating, max_lift, cargo_depth)


# merge with aggregated data
port_data_agg <- merge(port_data_agg, control_var, by.x = c("port_code", "year", "quarter"), by.y = c("port_code", "year", "quarter"), all.x = T) 

#port_data_agg$ships[port_data_agg$ships==0] <- 1

port_data_agg$quay_length[port_data_agg$quay_length==0] <- 100
port_data_agg$cranes[port_data_agg$cranes==0] <- 1

### Control variables for DID ####


controls <- c("cranes", "max_lift", "cargo_depth")
var_controls <- c("quay_length")

#controls <- c("cranes", "max_lift", "cargo_depth")
#var_controls <- c("quay_length")

```



```{r include=F, message=FALSE, warning=FALSE}

#### Estimation ot ATT with never treated as control group ####
attgt <- simple_staggered_did(yname = "container",            #outcome variable
                              tname = "period",                 #time variable
                              idname = "port_id",                    #id for units (should be port_code*sender)
                              gname = "group",                    #grouping variable for first treatment. Is zero for never treated units)
                              #xformula = c("ships"),
                              #varformula = c("ships"),
                              #control_group = "not_yet_treated",
                              data = port_data_agg)


attgt_cond <- simple_staggered_did(yname = "container",            #outcome variable
                              tname = "period",                 #time variable
                              idname = "port_id",                    #id for units (should be port_code*sender)
                              gname = "group",                    #grouping variable for first treatment. Is zero for never treated units)
                              xformula = controls,
                              varformula = var_controls,
                              #control_group = "not_yet_treated",
                              data = port_data_agg)



# putting together the results

results1 <- data.frame(Estimate = c("simple average", port_list$port[match(as.character(attgt$groupwise_treatment_effects$id), port_list$id)]),
                       ATT = c(attgt$average_treatment_effect$att,
                                   attgt$groupwise_treatment_effects$att),
                       lower = c(attgt$average_treatment_effect$att - attgt$average_treatment_effect$crit_val,
                                attgt$groupwise_treatment_effects$att - attgt$groupwise_treatment_effects$crit_val),
                       upper = c(attgt$average_treatment_effect$att + attgt$average_treatment_effect$crit_val,
                                  attgt$groupwise_treatment_effects$att + attgt$groupwise_treatment_effects$crit_val),
                       p_value = c(NA, attgt$groupwise_treatment_effects$pre_treatment_p_value))

results2 <- data.frame(Estimate = c("simple average", port_list$port[match(as.character(attgt_cond$groupwise_treatment_effects$id), port_list$id)]),
                       ATT_cond = c(attgt_cond$average_treatment_effect$att,
                                   attgt_cond$groupwise_treatment_effects$att),
                       lower_cond = c(attgt_cond$average_treatment_effect$att - attgt_cond$average_treatment_effect$crit_val,
                                attgt_cond$groupwise_treatment_effects$att - attgt_cond$groupwise_treatment_effects$crit_val),
                       upper_cond = c(attgt_cond$average_treatment_effect$att + attgt_cond$average_treatment_effect$crit_val,
                                  attgt_cond$groupwise_treatment_effects$att + attgt_cond$groupwise_treatment_effects$crit_val),
                        p_value_cond = c(NA, attgt_cond$groupwise_treatment_effects$pre_treatment_p_value))

results <- merge(results1, results2, by.x="Estimate", by.y="Estimate", all=T,  sort=F) %>%
               mutate(sig = (ATT>0 & lower>0) | (ATT<0 & upper<0),
                      sig_cond = (ATT_cond>0 & lower_cond>0) | (ATT_cond<0 & upper_cond<0),
                      pretest = (p_value < 0.05) & !is.na(p_value),
                      pretest_cond = (p_value_cond < 0.05) & !is.na(p_value_cond),
                      across(c("ATT","ATT_cond","lower","upper","lower_cond","upper_cond"), ~ format(round(.,digits=0), big.mark = ",")),

                      ATT = ifelse(sig==T, paste0(ATT, "*"), ATT),
                      ATT = ifelse(pretest==T, paste0(ATT, "^{x}"),ATT),
                      ATT_cond = ifelse(sig_cond==T, paste0(ATT_cond, "*"), ATT_cond),
                      ATT_cond = ifelse(pretest_cond==T, paste0(ATT_cond, "^{x}"),ATT_cond),
                      Conf = paste0("[",str_trim(lower),", ",str_trim(upper),"]"),
                      Conf_cond = paste0("[",str_trim(lower_cond),"; ",str_trim(upper_cond),"]"))
results <- results[c("Estimate", "ATT", "Conf", "ATT_cond", "Conf_cond")]



#### Estimate ATT with not-yet-treated control group ####
attgt_ny <- simple_staggered_did(yname = "container",            #outcome variable
                              tname = "period",                 #time variable
                              idname = "port_id",                    #id for units (should be port_code*sender)
                              gname = "group",                    #grouping variable for first treatment. Is zero for never treated units)
                              #xformula = c("ships"),
                              #varformula = c("ships"),
                              control_group = "not_yet_treated",
                              data = port_data_agg)


attgt_cond_ny <- simple_staggered_did(yname = "container",            #outcome variable
                              tname = "period",                 #time variable
                              idname = "port_id",                    #id for units (should be port_code*sender)
                              gname = "group",                    #grouping variable for first treatment. Is zero for never treated units)
                              xformula = controls,
                              varformula = var_controls,
                              control_group = "not_yet_treated",
                              data = port_data_agg)



# putting together the results

results_ny1 <- data.frame(Estimate = c("simple average", port_list$port[match(as.character(attgt_ny$groupwise_treatment_effects$id), port_list$id)]),
                       ATT_ny = c(attgt_ny$average_treatment_effect$att,
                                   attgt_ny$groupwise_treatment_effects$att),
                       lower_ny = c(attgt_ny$average_treatment_effect$att - attgt_ny$average_treatment_effect$crit_val,
                                attgt_ny$groupwise_treatment_effects$att - attgt_ny$groupwise_treatment_effects$crit_val),
                       upper_ny = c(attgt_ny$average_treatment_effect$att + attgt_ny$average_treatment_effect$crit_val,
                                  attgt_ny$groupwise_treatment_effects$att + attgt_ny$groupwise_treatment_effects$crit_val),
                       p_value_ny = c(NA, attgt_ny$groupwise_treatment_effects$pre_treatment_p_value))

results_ny2 <- data.frame(Estimate = c("simple average", port_list$port[match(as.character(attgt_cond_ny$groupwise_treatment_effects$id), port_list$id)]),
                       ATT_cond_ny = c(attgt_cond_ny$average_treatment_effect$att,
                                   attgt_cond_ny$groupwise_treatment_effects$att),
                       lower_cond_ny = c(attgt_cond_ny$average_treatment_effect$att - attgt_cond_ny$average_treatment_effect$crit_val,
                                attgt_cond_ny$groupwise_treatment_effects$att - attgt_cond_ny$groupwise_treatment_effects$crit_val),
                       upper_cond_ny = c(attgt_cond_ny$average_treatment_effect$att + attgt_cond_ny$average_treatment_effect$crit_val,
                                  attgt_cond_ny$groupwise_treatment_effects$att + attgt_cond_ny$groupwise_treatment_effects$crit_val),
                       p_value_cond_ny = c(NA, attgt_cond_ny$groupwise_treatment_effects$pre_treatment_p_value)) 

results_ny <- merge(results_ny1, results_ny2, by.x="Estimate", by.y="Estimate", all=T,  sort=F) %>%
               mutate(sig = (ATT_ny>0 & lower_ny>0) | (ATT_ny<0 & upper_ny<0),
                      sig_cond = (ATT_cond_ny>0 & lower_cond_ny>0) | (ATT_cond_ny<0 & upper_cond_ny<0),
                      pretest = (p_value_ny < 0.05) & !is.na(p_value_ny),
                      pretest_cond = (p_value_cond_ny < 0.05) & !is.na(p_value_cond_ny),
                      across(c("ATT_ny","ATT_cond_ny","lower_ny","upper_ny","lower_cond_ny","upper_cond_ny"), ~ format(round(.,digits=0), big.mark = ",")),
                      ATT_ny = ifelse(sig==T, paste0(ATT_ny, "*"), ATT_ny),
                      ATT_ny = ifelse(pretest==T, paste0(ATT_ny, "^{x}"),ATT_ny),
                      ATT_cond_ny = ifelse(sig_cond==T, paste0(ATT_cond_ny, "*"), ATT_cond_ny),
                      ATT_cond_ny = ifelse(pretest_cond==T, paste0(ATT_cond_ny, "^{x}"),ATT_cond_ny),
                      Conf_ny = paste0("[",str_trim(lower_ny),", ",str_trim(upper_ny),"]"),
                      Conf_cond_ny = paste0("[",str_trim(lower_cond_ny),"; ",str_trim(upper_cond_ny),"]"))
results_ny <- results_ny[c("Estimate", "ATT_cond_ny", "Conf_cond_ny")]

results <- merge(results, results_ny, by.x="Estimate", by.y="Estimate", sort=F)


```

```{r include=F}
# generating a gt table
results_gt <- gt(results, rowname_col = "Estimate") |>
              tab_header(
                title = "Aggregation of group time treatment effects, placebo treatment in period 80",
                subtitle = "Dependent variable: Total throughput with Lithuania in thousand TEU") |>
              cols_label(
                ATT="unconditional",
                ATT_cond="conditional",
                ATT_cond_ny="conditional",
                Conf="95% CI",
                Conf_cond="95% CI",
                Conf_cond_ny="95% CI"
              ) |>
              tab_spanner(label = "Never treated",
                          columns = c("ATT", "Conf", "ATT_cond", "Conf_cond")) |>
              tab_spanner(label = "Not yet treated",
                          columns = c("ATT_cond_ny", "Conf_cond_ny")) |>
              tab_row_group(label = "by group:",
                            rows = c(2:10)) |>
              row_group_order(groups = c(NA,"by group:")) |>
              sub_missing(columns = everything(),
                          rows = everything(),
                          missing_text = "") |>
              tab_footnote(footnote = "Stars indicate that the confidence interval does not include zero.") |>
              tab_options(stub.border.width = 0,
                          table_body.hlines.width = 0
              )

```

```{r include=TRUE, echo=FALSE}
results_gt
```

If we set the placebo treatments to earlier periods - pre-covid (90) and even longer ago (80 = 2016), we see that Gdynia traded with Latvia more than the average sample port of a similar size that was not treated. So there was no long term downwards trend kicking in before at least 2020, in contrary. 

# Difference in Differences: within-port variation in trade with partner ports in Lithuania as compared to partner ports in other countries

```{r include=FALSE, warning=FALSE, message=FALSE}
loadfonts(device = "win")
set.seed(100)

#load data
load("port_data_TEU_109.RData")

port_data_TEU <- filter(port_data_TEU, aggregate==0,
                                       stat_port==1,              #filter out non-stat ports (and geographic aggregates)
                                       TEU_yrly >=20000,          #filter out small ports
                                       port_code!="DEHAM") %>%      #exclude Hamburg
             mutate(s_china = ifelse(sender_code=="CN_X_HK" | sender_code=="HK", 1, 0)) %>%
             group_by(port_code, sender_code) %>%
             mutate(pair_id=cur_group_id()) %>% ungroup() %>%
             rename(container=total_TEU)

#exclude Rotterdam
port_data_TEU$container[port_data_TEU$port_code=="NLRTM" & between(port_data_TEU$year,2011,2020)]<-NA
port_data_agg$container[port_data_agg$port_code=="NLRTM" & between(port_data_agg$year,2011,2020)]<-NA


port_data_TEU <- port_data_TEU %>% 
  group_by(port_code) %>% 
  mutate(port_id = cur_group_id(),
         mean_container=mean(container))

# manipulate group variable to set treatment to September 2021 = period 100
port_data_TEU <- port_data_TEU %>%
  mutate(group = ifelse(group == 0, 0, 100))

port_list <- port_data_TEU %>% filter(group!=0, sender_code=="LT") %>% group_by(port) %>% summarize(id = mean(pair_id))

port_group_list <- port_data_TEU %>% filter(group!=0) %>% group_by(group) %>%
                                     summarize(ports = toString(unique(port)))
port_group_list$ports <- paste(port_group_list$ports,paste("(", port_group_list$group, ")", sep = ""))

port_data_TEU <- port_data_TEU %>%
  filter(group==100) %>%
  mutate(group=ifelse(sender_code!="LT", 0, 100))

# Add Control variables --------------------------------------------------------

# add terminal data to the data set
#terminal_data <- read_xlsx("port_manual_berth.xlsx")

load("berth_data.RData")
terminal_data <- berth_data
  
# add channel depth and cranes from World port index
wpi_data <- read.csv("WPI_data.csv")

# control variables
control_var <- merge(wpi_data, terminal_data, by.x="port_code",by.y="port_code", all = T) %>%
               select(port_code, 
                      year, quarter, quay_length, cranes, 
                      cranes_fixed, cranes_mobile, cranes_floating, max_lift, cargo_depth)


# merge with aggregated data
port_data_TEU <- merge(port_data_TEU, control_var, by.x = c("port_code", "year", "quarter"), by.y = c("port_code", "year", "quarter"), all.x = T) 

#port_data_agg$ships[port_data_agg$ships==0] <- 1

port_data_TEU$quay_length[port_data_agg$quay_length==0] <- 100
port_data_TEU$cranes[port_data_agg$cranes==0] <- 1

### Control variables for DID ####


controls <- c("cranes", "max_lift", "cargo_depth")
var_controls <- c("quay_length")

#controls <- c("cranes", "max_lift", "cargo_depth")
#var_controls <- c("quay_length")

```



```{r include=F, message=FALSE, warning=FALSE}

#### Estimation ot ATT with never treated as control group ####
attgt <- simple_staggered_did(yname = "container",            #outcome variable
                              tname = "period",                 #time variable
                              idname = "pair_id",                    #id for units (should be port_code*sender)
                              gname = "group",                    #grouping variable for first treatment. Is zero for never treated units)
                              #xformula = c("ships"),
                              #varformula = c("ships"),
                              #control_group = "not_yet_treated",
                              data = port_data_TEU)


attgt_cond <- simple_staggered_did(yname = "container",            #outcome variable
                              tname = "period",                 #time variable
                              idname = "pair_id",                    #id for units (should be port_code*sender)
                              gname = "group",                    #grouping variable for first treatment. Is zero for never treated units)
                              xformula = controls,
                              varformula = var_controls,
                              #control_group = "not_yet_treated",
                              data = port_data_TEU)



# putting together the results

results1 <- data.frame(Estimate = c("simple average", port_list$port[match(as.character(attgt$groupwise_treatment_effects$id), port_list$id)]),
                       ATT = c(attgt$average_treatment_effect$att,
                                   attgt$groupwise_treatment_effects$att),
                       lower = c(attgt$average_treatment_effect$att - attgt$average_treatment_effect$crit_val,
                                attgt$groupwise_treatment_effects$att - attgt$groupwise_treatment_effects$crit_val),
                       upper = c(attgt$average_treatment_effect$att + attgt$average_treatment_effect$crit_val,
                                  attgt$groupwise_treatment_effects$att + attgt$groupwise_treatment_effects$crit_val),
                       p_value = c(NA, attgt$groupwise_treatment_effects$pre_treatment_p_value))

results2 <- data.frame(Estimate = c("simple average", port_list$port[match(as.character(attgt_cond$groupwise_treatment_effects$id), port_list$id)]),
                       ATT_cond = c(attgt_cond$average_treatment_effect$att,
                                   attgt_cond$groupwise_treatment_effects$att),
                       lower_cond = c(attgt_cond$average_treatment_effect$att - attgt_cond$average_treatment_effect$crit_val,
                                attgt_cond$groupwise_treatment_effects$att - attgt_cond$groupwise_treatment_effects$crit_val),
                       upper_cond = c(attgt_cond$average_treatment_effect$att + attgt_cond$average_treatment_effect$crit_val,
                                  attgt_cond$groupwise_treatment_effects$att + attgt_cond$groupwise_treatment_effects$crit_val),
                        p_value_cond = c(NA, attgt_cond$groupwise_treatment_effects$pre_treatment_p_value))

results <- merge(results1, results2, by.x="Estimate", by.y="Estimate", all=T,  sort=F) %>%
               mutate(sig = (ATT>0 & lower>0) | (ATT<0 & upper<0),
                      sig_cond = (ATT_cond>0 & lower_cond>0) | (ATT_cond<0 & upper_cond<0),
                      pretest = (p_value < 0.05) & !is.na(p_value),
                      pretest_cond = (p_value_cond < 0.05) & !is.na(p_value_cond),
                      across(c("ATT","ATT_cond","lower","upper","lower_cond","upper_cond"), ~ format(round(.,digits=0), big.mark = ",")),

                      ATT = ifelse(sig==T, paste0(ATT, "*"), ATT),
                      ATT = ifelse(pretest==T, paste0(ATT, "^{x}"),ATT),
                      ATT_cond = ifelse(sig_cond==T, paste0(ATT_cond, "*"), ATT_cond),
                      ATT_cond = ifelse(pretest_cond==T, paste0(ATT_cond, "^{x}"),ATT_cond),
                      Conf = paste0("[",str_trim(lower),", ",str_trim(upper),"]"),
                      Conf_cond = paste0("[",str_trim(lower_cond),"; ",str_trim(upper_cond),"]"))
results <- results[c("Estimate", "ATT", "Conf", "ATT_cond", "Conf_cond")]



#### Estimate ATT with not-yet-treated control group ####
attgt_ny <- simple_staggered_did(yname = "container",            #outcome variable
                              tname = "period",                 #time variable
                              idname = "pair_id",                    #id for units (should be port_code*sender)
                              gname = "group",                    #grouping variable for first treatment. Is zero for never treated units)
                              #xformula = c("ships"),
                              #varformula = c("ships"),
                              control_group = "not_yet_treated",
                              data = port_data_TEU)


attgt_cond_ny <- simple_staggered_did(yname = "container",            #outcome variable
                              tname = "period",                 #time variable
                              idname = "pair_id",                    #id for units (should be port_code*sender)
                              gname = "group",                    #grouping variable for first treatment. Is zero for never treated units)
                              xformula = controls,
                              varformula = var_controls,
                              control_group = "not_yet_treated",
                              data = port_data_TEU)



# putting together the results

results_ny1 <- data.frame(Estimate = c("simple average", port_list$port[match(as.character(attgt_ny$groupwise_treatment_effects$id), port_list$id)]),
                       ATT_ny = c(attgt_ny$average_treatment_effect$att,
                                   attgt_ny$groupwise_treatment_effects$att),
                       lower_ny = c(attgt_ny$average_treatment_effect$att - attgt_ny$average_treatment_effect$crit_val,
                                attgt_ny$groupwise_treatment_effects$att - attgt_ny$groupwise_treatment_effects$crit_val),
                       upper_ny = c(attgt_ny$average_treatment_effect$att + attgt_ny$average_treatment_effect$crit_val,
                                  attgt_ny$groupwise_treatment_effects$att + attgt_ny$groupwise_treatment_effects$crit_val),
                       p_value_ny = c(NA, attgt_ny$groupwise_treatment_effects$pre_treatment_p_value))

results_ny2 <- data.frame(Estimate = c("simple average", port_list$port[match(as.character(attgt_cond_ny$groupwise_treatment_effects$id), port_list$id)]),
                       ATT_cond_ny = c(attgt_cond_ny$average_treatment_effect$att,
                                   attgt_cond_ny$groupwise_treatment_effects$att),
                       lower_cond_ny = c(attgt_cond_ny$average_treatment_effect$att - attgt_cond_ny$average_treatment_effect$crit_val,
                                attgt_cond_ny$groupwise_treatment_effects$att - attgt_cond_ny$groupwise_treatment_effects$crit_val),
                       upper_cond_ny = c(attgt_cond_ny$average_treatment_effect$att + attgt_cond_ny$average_treatment_effect$crit_val,
                                  attgt_cond_ny$groupwise_treatment_effects$att + attgt_cond_ny$groupwise_treatment_effects$crit_val),
                       p_value_cond_ny = c(NA, attgt_cond_ny$groupwise_treatment_effects$pre_treatment_p_value)) 

results_ny <- merge(results_ny1, results_ny2, by.x="Estimate", by.y="Estimate", all=T,  sort=F) %>%
               mutate(sig = (ATT_ny>0 & lower_ny>0) | (ATT_ny<0 & upper_ny<0),
                      sig_cond = (ATT_cond_ny>0 & lower_cond_ny>0) | (ATT_cond_ny<0 & upper_cond_ny<0),
                      pretest = (p_value_ny < 0.05) & !is.na(p_value_ny),
                      pretest_cond = (p_value_cond_ny < 0.05) & !is.na(p_value_cond_ny),
                      across(c("ATT_ny","ATT_cond_ny","lower_ny","upper_ny","lower_cond_ny","upper_cond_ny"), ~ format(round(.,digits=0), big.mark = ",")),
                      ATT_ny = ifelse(sig==T, paste0(ATT_ny, "*"), ATT_ny),
                      ATT_ny = ifelse(pretest==T, paste0(ATT_ny, "^{x}"),ATT_ny),
                      ATT_cond_ny = ifelse(sig_cond==T, paste0(ATT_cond_ny, "*"), ATT_cond_ny),
                      ATT_cond_ny = ifelse(pretest_cond==T, paste0(ATT_cond_ny, "^{x}"),ATT_cond_ny),
                      Conf_ny = paste0("[",str_trim(lower_ny),", ",str_trim(upper_ny),"]"),
                      Conf_cond_ny = paste0("[",str_trim(lower_cond_ny),"; ",str_trim(upper_cond_ny),"]"))
results_ny <- results_ny[c("Estimate", "ATT_cond_ny", "Conf_cond_ny")]

results <- merge(results, results_ny, by.x="Estimate", by.y="Estimate", sort=F)

```

```{r include=F}
# generating a gt table
results_gt <- gt(results, rowname_col = "Estimate") |>
              tab_header(
                title = "Aggregation of group time treatment effects, treatment in period 100",
                subtitle = "Within ports: Lithuania versus other countries") |>
              cols_label(
                ATT="unconditional",
                ATT_cond="conditional",
                ATT_cond_ny="conditional",
                Conf="95% CI",
                Conf_cond="95% CI",
                Conf_cond_ny="95% CI"
              ) |>
              tab_spanner(label = "Never treated",
                          columns = c("ATT", "Conf", "ATT_cond", "Conf_cond")) |>
              tab_spanner(label = "Not yet treated",
                          columns = c("ATT_cond_ny", "Conf_cond_ny")) |>
              tab_row_group(label = "by group:",
                            rows = c(2:10)) |>
              row_group_order(groups = c(NA,"by group:")) |>
              sub_missing(columns = everything(),
                          rows = everything(),
                          missing_text = "") |>
              tab_footnote(footnote = "Stars indicate that the confidence interval does not include zero.") |>
              tab_options(stub.border.width = 0,
                          table_body.hlines.width = 0
              )

```

```{r include=TRUE, echo=FALSE}
results_gt

```

The finding that Gdynia decreased trade with Lithuania after September 2021 more than other ports also holds when using  bilateral data and considering trade of the same port with other trading partners in the control group. Again, we find that trade with Lituania especially decreased for Gdynia. Gdynia decreased trade with Lithuania more than trade with its other trading partners. This holds for no other Chinese treated port.   

The finding may support the idea of China using its presence in the port of Gdynia as a policy tool to sanction Lithuania after it opened a diplomatic council. Yet, two alternative explanation remain which hinge on the impression that Gdynia is an important trade partner for Lithuanian ports: (1) covid hit all ports and the single ports were especially reducing trade with their most important trade partners, (2) the sanctioning of China doesn't take place at the port level but via an overall reduction of trade flows to Lithuania which is especially seen in reductions with the most important trade partners that may be Gdynia in this case.

# FE estimation

```{r include=FALSE, warning=FALSE, message=FALSE}
loadfonts(device = "win")
set.seed(100)

## aggregate estimation: only flows to Lithuania
#load data
load("port_data_TEU_109.RData")

port_data_TEU <- filter(port_data_TEU, aggregate==0,
                                       stat_port==1,              #filter out non-stat ports (and geographic aggregates)
                                       TEU_yrly >=20000,          #filter out small ports
                                       port_code!="DEHAM",        #exclude Hamburg due to very recent treatment
                                       sender_code=="LT") %>%      #include only containers to and from Lithuania
             mutate(s_china = ifelse(sender_code=="CN_X_HK" | sender_code=="HK", 1, 0)) %>%
             group_by(port_code, sender_code) %>%
             mutate(pair_id=cur_group_id()) %>% ungroup() %>%
             rename(container=total_TEU)


port_data_agg <- port_data_TEU %>% mutate(origin_china = ifelse(!is.na(container) & (sender_code=="CN_X_HK" | sender_code=="HK"),container,0)) %>%
                               group_by(port, port_code, period, reporter_code) %>% 
                               summarize(container = sum(container, na.rm=T),
                                         real_gdp_reporter = mean(real_gdp_reporter),
                                         origin_china = sum(origin_china),
                                         group = max(group),
                                         group_operation = max(group_operation),
                                         ownership_china = max(ownership_china),
                                         operation = max(operation),
                                         year = max(year),
                                         quarter = max(quarter)) %>% ungroup() %>%
                              group_by(port_code) %>% 
                              mutate(port_id = cur_group_id(),
                                     mean_container = mean(container, na.rm=T)) %>% ungroup()

# manipulate group variable to set treatment to September 2021 = period 100
port_data_agg <- port_data_agg %>%
  mutate(group = ifelse(group == 0, 0, 100))

port_data_agg$policy <- ifelse(port_data_agg$group == 100 & port_data_agg$period >= 100, 1, 0)
port_data_agg$Chinese_port <- ifelse(port_data_agg$group == 100, 1, 0)

#exclude Rotterdam
port_data_TEU$container[port_data_TEU$port_code=="NLRTM" & between(port_data_TEU$year,2011,2020)]<-NA
port_data_agg$container[port_data_agg$port_code=="NLRTM" & between(port_data_agg$year,2011,2020)]<-NA
```



```{r include=FALSE}
spec1 <- feols(container ~ policy  | port_code + period + reporter_code^period,
               data = port_data_agg)

# Extract the summary of the regression
summary_spec1 <- summary(spec1)

# Print the summary to see detailed results
print(summary_spec1)

# Extract coefficient estimates, R2, within R2, and number of observations
coef_table <- etable(spec1, tex = FALSE)

##################################

spec2 <- feols(container ~ policy | period + reporter_code^period,
               data = port_data_agg)

# Extract the summary of the regression
summary_spec2 <- summary(spec2)

# Print the summary to see detailed results
print(summary_spec2)

# Extract coefficient estimates, R2, within R2, and number of observations
coef_table2 <- etable(spec2, tex = FALSE)

######################
```

```{r include=FALSE, warning=FALSE, message=FALSE}
loadfonts(device = "win")
set.seed(100)

#load data
load("port_data_TEU_109.RData")

port_data_TEU <- filter(port_data_TEU, aggregate==0,
                                       stat_port==1,              #filter out non-stat ports (and geographic aggregates)
                                       TEU_yrly >=20000,          #filter out small ports
                                       port_code!="DEHAM") %>%      #exclude Hamburg
             mutate(s_china = ifelse(sender_code=="CN_X_HK" | sender_code=="HK", 1, 0)) %>%
             group_by(port_code, sender_code) %>%
             mutate(pair_id=cur_group_id()) %>% ungroup() %>%
             rename(container=total_TEU)

#exclude Rotterdam
port_data_TEU$container[port_data_TEU$port_code=="NLRTM" & between(port_data_TEU$year,2011,2020)]<-NA


port_data_TEU <- port_data_TEU %>% 
  group_by(port_code) %>% 
  mutate(port_id = cur_group_id(),
         mean_container=mean(container))

# manipulate group variable to set treatment to September 2021 = period 100
port_data_TEU <- port_data_TEU %>%
  mutate(group = ifelse(group == 0, 0, 100))

port_data_TEU$policy <- ifelse(port_data_TEU$group == 100 & port_data_TEU$period >= 100 & port_data_TEU$sender_code =="LT", 1, 0)

port_data_TEU_2 <- port_data_TEU %>%
  filter(group==100)

```

```{r include=FALSE}
spec3 <- feols(container ~ policy  | port_code^period + pair_id + sender_code^period,
               data = port_data_TEU)

# Extract the summary of the regression
summary_spec3 <- summary(spec3)

# Print the summary to see detailed results
print(summary_spec3)

# Extract coefficient estimates, R2, within R2, and number of observations
coef_table3 <- etable(spec3, tex = FALSE)



spec4 <- feols(container ~ policy | port_code^period + pair_id + sender_code^period,
               data = port_data_TEU_2)

# Extract the summary of the regression
summary_spec4 <- summary(spec4)

# Print the summary to see detailed results
print(summary_spec4)

# Extract coefficient estimates, R2, within R2, and number of observations
coef_table4 <- etable(spec4, tex = FALSE)



#set formatting names
setFixest_dict(c(ownership_china = "Ownership Chinese SOE",
                 s_china = "Origin China",
                 port_code = "Port",
                 period = "Period",
                 reporter_code = "Reporter",
                 sender = "Partner"))

#save table to latex 
cat(esttex(spec1, spec2, spec3, spec4,
                    depvar = FALSE,
                    style.df = style.df(depvar.title = "",
                                        fixef.title = "",
                                        fixef.suffix = "", yesNo = "yes"),
                    vcov = "twoway",
                    adjustbox = TRUE,
                    title = "Fixed effects Results",
                    label = "tab:fe_results"
                    #tex = TRUE
           ), file = file.path(getwd(),"results_tables","fe_results.tex"))

#prepare table for Markdownfile
fe_results<- etable(spec1, spec2, spec3, spec4,
                    depvar = FALSE,
                    style.df = style.df(depvar.title = "",
                                        fixef.title = "",
                                        stats.title = "",
                                        fixef.suffix = "", yesNo = "yes"),
                    vcov = "twoway")
colnames(fe_results)[1] <- "col1"

```

```{r include=TRUE, echo=FALSE}
print(fe_results)

```

We specify fixed effects estimation to check consistency with the previous DiD results on whether trade with Lithuania decreased after September 2021 for Chinese treated ports. We use the total sample in all specifications that includes all sample ports - treated and untreated.

In Spec1 and Spec2  we use only container flows to Lithuania (one observation per port per period). Spec3 and Spec4 use container flows to different partners (multiple observations per port per period). 

Policy turns 1 for all Chinese ports after period 100 (September 2021).

We cannot identify any global treatment effect for all ports. An identification of a Gdynia effect is not possible in this research design for it would only have 10 treated observations (Gdynia; Gdynia-LT in the periods 100-109). That is too few for treated observations for identification of an effect.


# Regression Discontinuity

## Checking validity of the Gdynia effect using RDD 

```{r include=TRUE, echo=FALSE}
loadfonts(device = "win")
set.seed(100)

load("port_data_TEU_109.RData")

port_data_TEU <- filter(port_data_TEU, aggregate==0,
                                       stat_port==1,              #filter out non-stat ports (and geographic aggregates)
                                       TEU_yrly >=20000,          #filter out small ports
                                       port_code!="DEHAM",        #exclude Hamburg due to very recent treatment
                                       sender_code=="LT",
                                        group!=0) %>%      #include only containers to and from Lithuania
             mutate(s_china = ifelse(sender_code=="CN_X_HK" | sender_code=="HK", 1, 0)) %>%
             group_by(port_code, sender_code) %>%
             mutate(pair_id=cur_group_id()) %>% ungroup() %>%
             rename(container=total_TEU)

#exclude Rotterdam
port_data_TEU$container[port_data_TEU$port_code=="NLRTM" & between(port_data_TEU$year,2011,2020)]<-NA

# drop observations before 90
#port_data_TEU <- port_data_TEU %>%
  #filter(period>=90)


# plot the container development 
ggplot(port_data_TEU, aes(period, container)) + 
  geom_point() +
  ggtitle("Container trade of all sample ports with Lithuania over time")

# Add Control variables --------------------------------------------------------

# add terminal data to the data set
#terminal_data <- read_xlsx("port_manual_berth.xlsx")

load("berth_data.RData")
terminal_data <- berth_data
  
# add channel depth and cranes from World port index
wpi_data <- read.csv("WPI_data.csv")

# control variables
control_var <- merge(wpi_data, terminal_data, by.x="port_code",by.y="port_code", all = T) %>%
               select(port_code, 
                      year, quarter, quay_length, cranes, 
                      cranes_fixed, cranes_mobile, cranes_floating, max_lift, cargo_depth)


# merge with aggregated data
port_data_TEU <- merge(port_data_TEU, control_var, by.x = c("port_code", "year", "quarter"), by.y = c("port_code", "year", "quarter"), all.x = T) 

#port_data_agg$ships[port_data_agg$ships==0] <- 1

port_data_TEU$quay_length[port_data_TEU$quay_length==0] <- 100
port_data_TEU$cranes[port_data_TEU$cranes==0] <- 1

# Create the treatment variable
port_data_TEU$treatment <- ifelse(port_data_TEU$period >= 100, 1, 0)


# (1) Gdynia data only
# (1.1) Fit the RD model - whole dataset

port_data_gdynia <- port_data_TEU %>%
  filter(port=="Gdynia")

# plot the container development of Gdynia to Lithuania
ggplot(port_data_gdynia, aes(period, container)) + 
  geom_point() +
  ggtitle("Container trade of Gdynia with Lithuania over time")


rd_model <- lm(container ~ treatment * I(period - 100) + cranes + max_lift + cargo_depth + quay_length, data = port_data_gdynia)


# Print the summary of the model
summary(rd_model)



# Plot the RD design
ggplot(port_data_gdynia, aes(x = period, y = container, color = factor(treatment))) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  labs(title = "Regression Discontinuity Design: Gdynia",
       x = "Period",
       y = "Container Throughput",
       color = "Treatment")
```

The underlying RDD model is:

\[
Y_{t} = \beta_0 + \beta_1 \text{RunningVar}_{t} + \beta_2 \text{Treated}_{t} + \beta_3 (\text{RunningVar}_{t} \times \text{Treated}_{t}) + \beta \mathbf{X}_{t} + \epsilon_{t}
\]

where:
- \( Y_{t} \) is the container throughput between Gdynia and Lithuania at time t.
- \( \text{RunningVar}_{t} \) is the running variable defined as the difference between period t and the cutoff period 100.
- \( \text{Treated}_{it} \) is is a dummy variable that takes the value 1 for periods greater than or equal to 100 and 0 otherwise.
- \( \mathbf{X}_{t} \) is a vector of control variables.
- \( \beta_0, \beta_1, \beta_2, \beta_3 \) are coefficients.
- \( \beta \) is a vector of coefficients for the control variables.
- \( \epsilon_{t} \) is the error term.

We do see a significant jump in the intercept and in the slope in the pre- as compared to the post- treatment period for Gdynia's trade with Lithuania.

## Gdynia placebo test: setting cutoff fat covid start period 94
```{r include=TRUE, echo=FALSE}
# (1.2) Placebo - different treatment period: 94 (covid)

# Create the alternative treatment variable
port_data_gdynia$treatment2 <- ifelse(port_data_gdynia$period >= 94, 1, 0)



rd_model_p <- lm(container ~ treatment2 * I(period - 94) + cranes + max_lift + cargo_depth + quay_length, data = port_data_gdynia)


# Print the summary of the model
summary(rd_model_p)



# Plot the RD design
ggplot(port_data_gdynia, aes(x = period, y = container, color = factor(treatment2))) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  labs(title = "Regression Discontinuity Design: Gdynia at placebo time 94 (covid)",
       x = "Period",
       y = "Container Throughput",
       color = "Treatment")
```

This change is not there if we set the cutoff / treatment at period 94 (this is when covid-19 hit) and estimate regressions to the left and to the right of the covid hit.


## Restricting Gdynia sample to 90-109 (around the cutoff at 100)
```{r include=TRUE, echo=FALSE}



# (1.3) Restricted dataset from 90-109

port_data_gdynia_90 <- port_data_gdynia %>%
  filter(period>=90)

rd_model_rs <- lm(container ~ treatment * I(period - 100) + cranes + max_lift + cargo_depth + quay_length, data = port_data_gdynia_90)


# Print the summary of the model
summary(rd_model_rs)

# Plot the RD design
ggplot(port_data_gdynia_90, aes(x = period, y = container, color = factor(treatment))) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  labs(title = "Regression Discontinuity Design: Gdynia restricted sample since period 90",
       x = "Period",
       y = "Container Throughput",
       color = "Treatment")
```

For calculating the regression line in the previous RDD estimations, we used the full sample. If we restrict our sample around the cutting point +/- 9 periods, our effect remains of the same size and directions but loses significance. This is probalby due to the small number of observations on either side of the cutoff.

## Using all Chinese-treated ports not just Gdynia

```{r include=TRUE, echo=FALSE}
# (2a) all ports version with port dummies
# Fit the RD model

# plot the container development of all treated ports to Lithuania
ggplot(port_data_TEU, aes(period, container)) + 
  geom_point()

rd_model2 <- lm(container ~ treatment * I(period - 100) + cranes + max_lift + cargo_depth + quay_length + factor(port_code), data = port_data_TEU)


rd_model2 <- lm(container ~ treatment * I(period - 100) + factor(port_code), data = port_data_TEU)

# Print the summary of the model
summary(rd_model2)

# Plot the RD design
ggplot(port_data_TEU, aes(x = period, y = container, color = factor(treatment))) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  labs(title = "Regression Discontinuity Design: all ports",
       x = "Period",
       y = "Container Throughput",
       color = "Treatment")
```

When we include all treated ports in the sample and check their trade with Lithuania, we do not find a significant negative effect any more. The direction and size of the intercept jump and slope change remains the same as for the restricted sample on Gdynia. 

The underlying model that was estimated takes care of changes in the port infrastructure and includes port-dummies. More precisely, we estimate :


\[
Y_{it} = \beta_0 + \beta_1 \text{RunningVar}_{it} + \beta_2 \text{Treated}_{it} + \beta_3 (\text{RunningVar}_{it} \times \text{Treated}_{it}) + \beta \mathbf{X}_{it} + \alpha_i + \epsilon_{it}
\]

where:
- \( Y_{it} \) is the the container throughput between China trated port i and Lithuania at time t.
- \( \text{RunningVar}_{it} \) is the running variable defined as the difference between period t and the cutoff period 100.
- \( \text{Treated}_{it} \) is a dummy variable that takes the value 1 for periods greater than or equal to 100 and 0 otherwise.
- \( \mathbf{X}_{it} \) is a vector of control variables.
- \( \beta_0, \beta_1, \beta_2, \beta_3 \) are coefficients.
- \( \beta \) is a vector of coefficients for the control variables.
- \( \alpha_i \) is the individual-specific effect.
- \( \epsilon_{it} \) is the error term.

The coefficient \(\beta_2\) captures the average treatment effect of the policy change at the cutoff period (level effect), while \(\beta_3\) captures how this effect varies with the running variable (slope effect).

## RDD with FE using the sample with all Chinese ports

```{r include=TRUE, echo=FALSE}

# (2b) all ports version formulating a FE model with plm

port_data_TEU <- pdata.frame(port_data_TEU, index = c("port_code", "period"))

# Convert period to numeric if it's not already
port_data_TEU$period <- as.numeric(as.character(port_data_TEU$period))

# Ensure port_code is a factor
port_data_TEU$port_code <- as.factor(port_data_TEU$port_code)

# Fit the RD model with port fixed effects using plm
rd_model3 <- plm(container ~ treatment * I(period - 100) + cranes + max_lift + cargo_depth + quay_length, 
                data = port_data_TEU, model = "within", effect = "individual")

summary(rd_model3)
```

When using a panel RDD estimator that uses within variation only, we get similar estimates.

A positive jump of the intercept at 100 and then a downwards sloping line that is however insignificant. Suggesting that there is no overall effect of all treated ports but that it is rather Gdynia that drives the outcome.

## Checking other ports trade development with Lituania


```{r include=TRUE, echo=FALSE}
# placebo: other port (Tallinn, Riga, Oslo)

# set up dataset again

loadfonts(device = "win")
set.seed(100)

load("port_data_TEU_109.RData")

port_data_TEU <- filter(port_data_TEU, aggregate==0,
                                       stat_port==1,              #filter out non-stat ports (and geographic aggregates)
                                       TEU_yrly >=20000,          #filter out small ports
                                       port_code!="DEHAM",        #exclude Hamburg due to very recent treatment
                                       sender_code=="LT") %>%      #include only containers to and from Lithuania
             mutate(s_china = ifelse(sender_code=="CN_X_HK" | sender_code=="HK", 1, 0)) %>%
             group_by(port_code, sender_code) %>%
             mutate(pair_id=cur_group_id()) %>% ungroup() %>%
             rename(container=total_TEU)

#exclude Rotterdam
port_data_TEU$container[port_data_TEU$port_code=="NLRTM" & between(port_data_TEU$year,2011,2020)]<-NA

# Add Control variables --------------------------------------------------------

# add terminal data to the data set
#terminal_data <- read_xlsx("port_manual_berth.xlsx")

load("berth_data.RData")
terminal_data <- berth_data
  
# add channel depth and cranes from World port index
wpi_data <- read.csv("WPI_data.csv")

# control variables
control_var <- merge(wpi_data, terminal_data, by.x="port_code",by.y="port_code", all = T) %>%
               select(port_code, 
                      year, quarter, quay_length, cranes, 
                      cranes_fixed, cranes_mobile, cranes_floating, max_lift, cargo_depth)


# merge with aggregated data
port_data_TEU <- merge(port_data_TEU, control_var, by.x = c("port_code", "year", "quarter"), by.y = c("port_code", "year", "quarter"), all.x = T) 

#port_data_agg$ships[port_data_agg$ships==0] <- 1

port_data_TEU$quay_length[port_data_TEU$quay_length==0] <- 100
port_data_TEU$cranes[port_data_TEU$cranes==0] <- 1

# Create the treatment variable
port_data_TEU$treatment <- ifelse(port_data_TEU$period >= 100, 1, 0)

# Fit the RD model Tallinn

port_data_Tallinn <- port_data_TEU %>%
  filter(port=="Tallinn")

# plot the container development of Tallinn to Lithuania
ggplot(port_data_Tallinn, aes(period, container)) + 
  geom_point() +
  ggtitle("Container trade of Tallinn with Lithuania over time")


rd_model4 <- lm(container ~ treatment * I(period - 100) + cranes + max_lift + cargo_depth + quay_length, data = port_data_Tallinn)


# Print the summary of the model
summary(rd_model4)


# Plot the RD design
ggplot(port_data_Tallinn, aes(x = period, y = container, color = factor(treatment))) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  labs(title = "Regression Discontinuity Design: Tallinn",
       x = "Period",
       y = "Container Throughput",
       color = "Treatment")

# Fit the RD model Riga
port_data_Riga <- port_data_TEU %>%
  filter(port=="Riga")

# plot the container development of Riga to Lithuania
ggplot(port_data_Riga, aes(period, container)) + 
  geom_point() + 
  ggtitle("Container trade of Riga with Lithuania over time")


rd_model5 <- lm(container ~ treatment * I(period - 100) + cranes + max_lift + cargo_depth + quay_length, data = port_data_Riga)


# Print the summary of the model
summary(rd_model5)


# Plot the RD design
ggplot(port_data_Riga, aes(x = period, y = container, color = factor(treatment))) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  labs(title = "Regression Discontinuity Design: Riga",
       x = "Period",
       y = "Container Throughput",
       color = "Treatment")


# Fit the RD model Oslo
port_data_Oslo <- port_data_TEU %>%
  filter(port=="Oslo")

# plot the container development of Oslo to Lithuania
ggplot(port_data_Oslo, aes(period, container)) + 
  geom_point() +
  ggtitle("Container trade of Oslo with Lithuania over time")


rd_model6 <- lm(container ~ treatment * I(period - 100) + cranes + max_lift + cargo_depth + quay_length, data = port_data_Oslo)


# Print the summary of the model
summary(rd_model6)


# Plot the RD design
ggplot(port_data_Oslo, aes(x = period, y = container, color = factor(treatment))) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  labs(title = "Regression Discontinuity Design: Oslo",
       x = "Period",
       y = "Container Throughput",
       color = "Treatment")

```

Tallinn, Riga, Oslo are geographically close and have a similar average trade volume with Lithuania as compared to Gdynia. 

Tallinn shows a similar behavior as Gdynia. Its post-treatment slope coefficient is negative too but insignificant. Oslo too is similar and shows a significant negative slope after treatement. Riga behaves very differently to Gdynia and Oslo. 

The results alltogether suggest that Gdynia did reduce its trade to Lithuania after the diplomatic dispute between China and Lithuania. It is the only Chinese-treated port to do so. It also behaves differently than the average comparable non-treated port in the sample but it does not behave differently to single other comparable ports in the region, especially Tallinn, and Oslo. 

The policy effect may be one that goes without port ownership by Chinese but by simply restricting trade with Lithuania that hits those ports more that are heavily involved in trade with Lithuanian ports.

# Event Study 

Upcoming on July 18/19 ...